### 2023-06-30 Article.tex
Filerne er tilføjet nu

\documentclass[sigplan,screen]{acmart}
\input{preamble}
%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[GPCE 2023]{Make sure to enter the correct
  conference title from your rights confirmation emai}{July 07--07,
  2023}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{A generalised editor calculus (Short Paper)}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

%%\orcid{1234-5678-9012}

%\author{Andreas Tor Mortensen}
%\email{atmo20@student.aau.dk}

%\author{Benjamin Bennetzen}
%\email{bbenne20@student.aau.dk}

%\author{Nikolaj Rossander Kristensen}
%\email{nrkr20@student.aau.dk}

%\author{Peter Buus Steffensen}
%%\authornotemark[1]
%\email{psteff19@student.aau.dk
%}

%\affiliation{%
%  \institution{Department of Computer Science at Aalborg University}
%  \streetaddress{Selma Lagerlöfsvej 300}
%  \city{Aalborg}
%  \country{Denmark}
%  \postcode{9220}
%}

%\author{Hans Huttel}
%\email{hans@cs.aau.dk}
%\affiliation{%
%  \institution{Department of Computer Science at Aalborg University}
%  \streetaddress{Selma Lagerlöfsvej 300}
 % \city{Aalborg}
 % \country{Denmark}
 % \postcode{9220}
%}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{AAU Bachelor}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
In this paper, we present a generalization of a syntax-directed editor calculus, which can be used to instantiate a specialized syntax-directed editor for any language, given by some abstract syntax. The editor calculus guarantees the absence of syntactical errors while allowing incomplete programs. The generalized editor calculus is then encoded into a simply typed lambda calculus, extended with pairs, booleans, pattern matching and fixed points.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Syntax-directed editor, Generalisation, lambda calculus}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%\begin{teaserfigure}
 % \includegraphics[width=\textwidth]{Images/windmill.jpg}
 % \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
This paper is on the topic of syntax-directed editors. 
Most current examples of syntax-directed editor calculi, such as the ones presented by \cite{hazelnut} and \cite{type_safe_structure_editor}, are typically tailored specifically towards editing abstract syntax trees from an applied $\lambda$-calculus. 
This fact alone limits the usable scope of these calculi, since they would be rendered useless given any other form of abstract syntax.

Furthermore, these calculi do not account for binding mechanisms in abstract syntax. Many of the editor calculi possess the ability to interleave construction and evaluation of programs, however, a consequence of this is a complex type system.

In this paper we propose a generalized editor calculus that can be used to create a syntax-directed editor calculus for any abstract syntax. 
Our editor calculus is inspired by higher-order abstract syntax presented in \cite{hoas}, and it will take binding-mechanisms into account. We have encoded our generalized editor calculus into a simply-typed lambda calculus extended with pairs, booleans, pattern matching and fixed-points.

%Copy paste fra her
\section{Generalized Editor Calculus}\label{sec:general_editor}
We will now present a generalization of the syntax-directed editor calculus by \cite{type_safe_structure_editor}. 
This specific editor calculus was designed with the language of an applied $\lambda$-calculus in mind. 
%Instead, by using the notion of abstract syntax as presented in \cite{sec:pre_abstract_syntax} we generalize the concepts of the editor calculus to account for any language given by some abstract syntax.

\subsection{Syntax}

\subsubsection{Abstract Syntax}
Firstly, we introduce the general representation of some abstract syntax that will be the focus of the editor calculus. We assume the abstract syntax is given by a set of sorts $\SORT$, an arity-indexed family of operators $\OP$ and a sort-indexed family of variables $\VAR$, such as presented in \cite{harper_foundations}. For the full definitions we refer to the full version of the paper.

The notion of cursors and holes is central to the idea of an editor calculus. Whereas the calculus in \cite{type_safe_structure_editor} only has one cursor and hole term, in the general case we need a cursor and hole operator for every sort in the abstract syntax.
With the set of sorts $\SORT$, the extended family of operators $\OP$ and family of variables $\VAR$ we can then define the sort-indexed family of abstract binding trees (refered to as "abt") $\ABT{}$ as the smallest family that satisfies the conditions of $\ABT{}$'s in \cite{harper_foundations}.

\begin{example}\label{ex:abstract_syntax}
    %To illustrate how the generalized editor calculus could be instantiated for some language, we will specialize an editor calculus as we proceed through this section for the abstract syntax shown in \cref{fig:example_abstract_syntax}.
    To illustrate how the generalized editor calculus could be instantiated for some language, we will through our examples specialize an editor calculus. The abstract syntax for our specialization language is shown in \cref{fig:example_abstract_syntax}.
    \federeboks{R C L L L}{
        \text{Sort} && \text{Term} & \text{Operator} & \text{Arity} \\
        s & ::= & \text{let $x = e$ in $s$} & \text{let} & (e, e.s)s \\
        & | & e & \text{exp} & (e)s \\
        e & ::= & e_1 + e_2 & \text{plus} & (e,e)e \\
        & | & n & \text{num}[n] & ()e \\
        & | & x & \text{var}[x] & ()e
    }{\caption{Abstract syntax of arithmetic expressions and local declarations}\label{fig:example_abstract_syntax}}
    The syntax in \cref{fig:example_abstract_syntax} describes a very simple language of statements and expressions. For example, using the abstract syntax we could construct the statement
    \begin{equation*}
        \text{let $x=5$ in let $y=10$ in $x+y$}
    \end{equation*}
    which, using operators, would be represented as the abt
    \begin{equation*}
        \text{let}(5;x.\text{let}(10;y.\text{exp}(\text{plus}(x;y))))
    \end{equation*}

    Given the abstract syntax in \cref{fig:example_abstract_syntax} we can extend it with holes and cursors. This gives us the new operators shown in \cref{fig:example_holes_and_cursors}.
    \federeboks{R C L L L}{
        \text{Sort} && \text{Term} & \text{Operator} & \text{Arity} \\
        s & ::= & [s] & \text{cursor}_s & (s)s \\
        & | & \hole{s} & \text{hole}_s & ()s \\
        e & ::= & [e] & \text{cursor}_e & (e)e \\
        & | & \hole{e} & \text{hole}_e & ()e
    }{\caption{Introduction of cursors and holes to the abstract syntax in \cref{fig:example_abstract_syntax}}\label{fig:example_holes_and_cursors}}
    With this extension we can now represent statements with holes and cursors as abt's. For example the statement
    \begin{equation*}
        \text{let $x=[5]$ in $x+\hole{e}$}
    \end{equation*}
    would be represented as the abt
    \begin{equation*}
        \text{let}(\text{cursor}_e(5); x.\text{exp}(\text{plus}(x;\text{hole}_e)))    
    \end{equation*}
\end{example}

\subsubsection{Editor Calculus}
The abstract syntax of the generalized editor calculus is shown in \cref{fig:editor_calculus_syntax}. It closely resembles that of the calculus presented by \cite{type_safe_structure_editor}.
\federeboks{R C L}{
    E & ::= & \pi.E \ | \ \phi => E_1|E_2 \ | \ E_1 \ggg E_2 \ | \ \text{rec} \ x.E \ | \ x \ | \ \text{nil} \\
    \pi & ::= & \text{child} \ n \ | \ \text{parent} \ | \ \{o\} \\
    \phi & ::= & \neg\phi \ | \ \phi_1 \wedge \phi_2 \ | \ \phi_1 \vee \phi_2 \ | \ @o \ | \ \Diamond o \ | \ \square o
}{\caption{Abstract syntax of general editor calculus}\label{fig:editor_calculus_syntax}}
Editor expressions $E \in Edt$, where $Edt$ is the sort of editor expressions, describes the behaviour of the editor calculus. The semantics behind the different constructs are defined in section \cref{sec:Semantics}.

Evaluation of abt's, which \cite{type_safe_structure_editor} introduces through a $eval$ construct, is not included in our generalized version. This omission is due to evaluation being entirely dependent on the dynamics of the language in focus, and not the static and structural properties the abstract syntax describes. To define the concept of evaluation in a generalized manner, it would be necessary to devise a formal and concise approach for specifying the semantics of operators.

\begin{example}\label{ex:editor_calculus}
    Expanding upon \cref{ex:abstract_syntax}, we can specialize the abstract syntax of the general editor calculus presented in \cref{fig:editor_calculus_syntax} for our simple language by defining the set of operators $o$ ranges over. In this case we have that
    \begin{equation*}
        o \in \{ \text{let}, \text{exp}, \text{hole}_s, \text{plus}, \text{num}[n], \text{var}[x], \text{hole}_e \}
    \end{equation*}
    allowing us to write editor expressions such as $(@\text{hole}_e => \{\text{plus}\}.\text{nil}|\text{nil})$, which would substitute the current tree encapsulated by the cursor with the $plus$ operator, if the cursor is at the operator $\text{hole}_e$.
\end{example}

\subsubsection{Cursor Context}
In editor expressions, we frequently refer to performing actions on, or with, the abt that is currently encapsulated by the cursor. To support this notion we introduce cursor contexts $C$, inspired by the zipper data structure by \cite{huet_zipper}, as a way to locate the cursor in an abt. Central to this idea is the cursor context $\ctxhole$, or context hole, which specifies the subtree within which the cursor must reside, either as the root, or one of the first children.  For the full definition of cursor contexts, we refer to the full version of the paper, or to \cite{type_safe_structure_editor}, as they have a very similar construct.

%To ensure that the cursor can only reside somewhere within the subtree specified by the context hole, we need to introduce some notion of cursorless abt's, written as $\hat{s}$. Using the abstract syntax, given by $\SORT$, $\OP$ and $\VAR$, we define a new corresponding abstract syntax which can only build abt's without cursors.

%The abstract syntax of cursor contexts are defined according to \cref{def:cursor_contexts}. Using cursor contexts we can interpret any sort of abt $a \in \ABT{}$ as the cursor context $C[a']$, where the abt $a'$ is substituted for the context hole $\ctxhole$. Due to \cref{def:cursor_contexts} this also means that the cursor can only reside within $a'$, since it cannot be anywhere in $C$.

%\begin{definition}[Cursor Contexts]\normalfont
%    \label{def:cursor_contexts}
 %   The abstract syntax of cursor contexts is defined as an extension of the syntax for cursorless abt's. It is given by
%    \begin{enumerate}
%        \item The sorts $\SORT^C = \SORTc \cup \{ C \}$
%        \item The family of operators $\OP^C = \OPc$ extended with the $\ctxhole$ operator of arity $()C$.
 %       \item Furthermore, for every operator $\hat{o} \in \OPc$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_n}.\hat{s}_n)\hat{s}$ and for every $1 \leq i \leq n$ we also add the operator $o^C_i$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_i}.C, \dots, \Vec{\hat{s}_n}.\hat{s}_n)C$ to $\OP^C$. These represents that $\ctxhole$ must be somewhere in subtree $i$ of operator $\hat{o}$.
%        \item The family of variables $\VAR^C = \VARc$
%    \end{enumerate}
%\end{definition}

%\begin{example}\label{ex:cursor_contexts}
%    Continuing from \cref{ex:editor_calculus}, we will now extend the editor calculus for our simple language with cursorless abt's and cursor contexts.

 %   Firstly, we define the abstract syntax of abt's without cursors, the result of which is shown in \cref{fig:example_cursorless_trees}. For brevity the syntax is presented as formation rules, since translating these into the notion of operators, while trivial, is not a very concise way of presenting the abstract syntax.
 %   \federeboks{R C l}{
  %      \hat{s} & ::= & let $x = \hat{e}$ in $\hat{s}$ | %$\hat{e}$ | $\hole{s}$ \\
  %      \hat{e} & ::= & $\hat{e}_1 + \hat{e}_2$ | n | x | $\hole{e}$
 %   }{\caption{Abstract syntax of abt's without cursors}\label{fig:example_cursorless_trees}}
 %   Secondly, using the syntax of cursorless abt's we can define the abstract syntax of cursor contexts according to \cref{def:cursor_contexts}. This is shown in \cref{fig:example_cursor_contexts}.
 %   \federeboks{R C l}{
 %       C & ::= & let $x = C$ in $\hat{s}$ | let $x = \hat{e}$ in $C$ \\
  %      & | & $C + \hat{e}_2$ | $\hat{e}_1 + C$ | $\ctxhole$
 %   }{\caption{Abstract syntax of cursor contexts}\label{fig:example_cursor_contexts}}
%\end{example}

Although a cursor context accurately specifies the subtree where the cursor must reside, it does not inherently guarantee the presence of only one cursor within that subtree. Therefore, it becomes necessary to introduce the concept of well-formed abt's. We will do this informally, where a well-formed abt is simply an abt containing exactly one cursor. This is guarantied when an abt can be interpreted as $C[\dot{a}]$, meaning the context C, where $\ctxhole$ is substituted with a tree $\dot{a}$, where the cursor-operator is either the root $\dot{a}$, or one of the children of the root.

%\begin{definition}[Well-formed trees]\normalfont
%    \label{def:well_formed_trees}
%    Firstly, we define an abstract syntax that describes abt's with exactly one cursor either as the root or as an argument of the root. The syntax is given by
 %   \begin{enumerate}
 %       \item The sorts $\SORTcc = \SORTc \cup \{ \dot{s} \}_{s \in \SORT}$.
 %       \item The family of operators $\OPcc = \OPc$ extended with an operator of arity $(\hat{s})\dot{s}$ for every $\hat{s} \in \SORTc$. These represent that the cursor encapsulates the root of a cursorless abt of sort $\hat{s}$.
   %     \item Furthermore, for every operator $\hat{o} \in \OPc$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_n}.\hat{s}_n)\hat{s}$ and for every $1 \leq i \leq n$ we also add the operator $\dot{o}_i$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_i}.\hat{s}_i, \dots, \Vec{\hat{s}_n}.\hat{s}_n)\dot{s}$ to $\OPcc$. These represent that the cursor encapsulates the cursorless subtree $i$ of operator $\hat{o}$.
%        \item The family of variables $\VARcc = \VARc$.
 %   \end{enumerate}
%    Given this, a well-formed abt $a \in \ABT{}$ is any abt that can be interpreted as $C[\dot{a}]$, since $C$ cannot contain any cursors and $\dot{a}$ contains exactly one.
%\end{definition}

\begin{example}\label{ex:well_formed_trees}
   We will now show the extension of \cref{fig:example_abstract_syntax}, where cursors are included, as shown in \cref{fig:example_well_formed_trees}. We note that $\hat{s}$ and $\hat{e}$ are trees without cursors.
    \federeboks{R C l}{
        \dot{s} & ::= & let $x = [\hat{e}]$ in $\hat{s}$ | let $x = \hat{e}$ in $[\hat{s}]$ | $[\hat{e}]$ | $[\hat{s}]$ \\
        \dot{e} & ::= & $[\hat{e}_1] + \hat{e}_2$ | $\hat{e}_1 + [\hat{e}_2]$ | $[\hat{e}]$
    }{\caption{Abstract syntax of abt's with exactly one cursor}\label{fig:example_well_formed_trees}}
    Given the syntax in \cref{fig:example_well_formed_trees} and the previously introduced cursor context, we can now determine the well-formedness of any abt in our language. For example the statement
    \begin{equation*}
        \text{let $x=[\hole{e}]$ in $x+5$}
    \end{equation*}
    is well-formed since it can be written as $C[\dot{e}]$ where
    \begin{align*}
        C &= \text{let $x=\ctxhole$ in $x+5$} \\
        \dot{e}&= [\hole{e}]
    \end{align*}
    It should be noted that there often are two valid interpretations of a tree $a$ as $C[\dot{a}]$. In this example, the other valid interpretation would be:
    \begin{align*}
        C' &= \ctxhole \\
        \dot{e}' &= \text{let $x=[\hole{e}]$ in $x+5$}
    \end{align*}
\end{example}

\subsection{Semantics}
In this section we present the transition systems and the general forms of transition rules defining these systems, based on the generalized syntax of the editor calculus previously presented.

\subsubsection{Editor Expressions} \label{sec:Semantics}
The labelled transition system for editor expressions is defined as $(Edt \times \ABT{}, Apc \cup \{\epsilon\}, \Rightarrow)$. Transitions are of the form $\lr{E, a} \ltrans{\alpha} \lr{E', a'}$ where the editor expression $E \in Edt$ is closed and the abt $a \in \ABT{}$ is well-formed. The labels of the transitions, $\alpha$, are the atomic prefix commands $Apc$ and the silent transition $\epsilon$. The transition relation $\Rightarrow$ is defined by the rules shown in \cref{fig:editor_expression_rules}. These facilitate ways of composing editor expressions together with conditions, sequential composition and recursion. The means of traversing and modifying an abt are provided by the prefixed expression $\pi.E$, which evaluates $\pi$ before continuing with $E$. The conditional expression $\phi => E_1|E_2$ reduces to $E_1$ if $\phi$ is satisfied and $E_2$ otherwise. The sequential expression $E_1 \ggg E_2$ evaluates $E_2$ only once $E_1$ has been reduced to $nil$. The recursive expression $rec \ x.E$ binds the recursion variable $x$ in $E$, which can then be used to recursively iterate an expression.

Unlike the editor calculus by \cite{type_safe_structure_editor}, we do not utilize any form of structural congruence on editor expressions to define the semantics of trivial sequential composition or recursion. Instead we simply define these notions with their own transition rule (Seq-Trivial) and (Recursion)
\federeboks{c}{
    \infeshort{Cond-1}
            {a |= \phi}
            {\lr{\phi => E_1 | E_2, C[a]} \ltrans{\epsilon} \lr{E_1, C[a]}}
    \\\\
    \infeshort{Cond-2}
        {a \nvDash \phi}
        {\lr{\phi => E_1 | E_2, C[a]} \ltrans{\epsilon} \lr{E_2, C[a]}}
    \\\\
    \infeshort{Seq}
        {\lr{E_1, a} \ltrans{\alpha} \lr{E_1',a'}}
        {\lr{E_1 \ggg E_2, a} \ltrans{\alpha} \lr{E_1' \ggg E_2,a'}}
    \\\\
    \infeshort{Seq-Trivial}
        {}
        {\lr{\text{nil} \ggg E_2, a} \ltrans{\epsilon} \lr{E_2,a}}    
    \\\\
    \infeshort{Recursion}
        {}
        {\lr{\text{rec } x.E,a} \ltrans{\epsilon} \lr{E[x:=rec \ x.E], a}}
    \\\\
    \infeshort{Context}
        {a \ltrans{\pi} a'}
        {\lr{\pi.E, C[a]} \ltrans{\pi} \lr{E, C[a']}}
}{\caption{Reduction rules for editor expressions}\label{fig:editor_expression_rules}}
.

\subsubsection{Substitution and cursor movement}
The labelled transition system for substitutions and cursor movement is defined as $(\ABT{}, Apc, \Rightarrow)$. This system describes the semantics of modifying the well-formed abt $a$ encapsulated by the cursor, by either substituting it with an operator or moving the cursor up or down the tree. Transitions are therefore of the form $a \ltrans{\pi} a'$, where $a, a' \in \ABT{}$ and $\pi \in Apc$.

For substitution we define the transition rule for every label $\{o\} \in Apc$, which substitutes the abt currently encapuslated by the cursor with $o$. The general case of this rule is shown in \cref{fig:substitution_rules}. The side-condition ensures that we can only substitute operators of sort $s$ with abt's of sort $s$.
\federeboks{c}{
    \infeshort{Ins}
        {}
        {[\hat{a}]  \ltrans{\{o\}} [o(\Vec{x_1}.\hole{{s_1}}; \dots ; \Vec{x_n}.\hole{{s_n}})]}
        $\begin{array}{l}
         \hat{a}, o \in \ABT{}_s % , \\ \text{ where $s$ is the sort of $o$}
        \end{array}$
}{\caption{General form of reduction rule for substitution}\label{fig:substitution_rules}}

For cursor movement we define two transition rules "child  $i$" and "parent" for every operator $o$ of arity $(\Vec{s_1}.s_1, \dots,$ $ \Vec{s_n}.s_n)s$ and for every $1 \leq i \leq n$. These rules, shown in \cref{fig:cursor_movement_rules}, facilitate cursor movement from the parent operator to child $i$ and from child $i$ back to the parent operator, respectively.
\federeboks{c}{
    \infeshort{Child-i}
        {}
        {[\hat{o}(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_n}.\hat{a}_n)] \ltrans{\text{child} \ i}} \\ 
        \quad \quad \quad \quad \quad $o(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_i}.[a_i]; \dots; \Vec{x_n}.\hat{a}_n)$
    \\\\
    \infeshort{Parent}
        {}
        {o(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_i}.[\hat{a}_i]; \dots; \Vec{x_n}.\hat{a}_n) \ltrans{\text{parent}}} \\
        \quad \quad \quad \quad \quad $[\hat{o}(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_n}.\hat{a}_n)]$
}{\caption{General form of reduction rules for cursor movement}\label{fig:cursor_movement_rules}}

\begin{example}\label{ex:substitution_rules}
    Continuing from \cref{ex:well_formed_trees}, we now define the semantics of the editor calculus for our simple language. We do not show the transition rules for editor expressions, since these are equivalent to the ones presented in \cref{fig:editor_expression_rules}.

    The transition system for cursor movement and substitution is defined with respect to the operators $o$ in our language. Based on the general form of substitution rules shown in \cref{fig:substitution_rules}, we define a specialized rule for substituting the tree encapsulated by the cursor for every operator $o$. A selection of these are shown in \cref{fig:example_substitution_rules}.
    \federeboks{c}{
        \infelong{let}
            {}
            {[\hat{a}]  \ltrans{\{\text{let}\}} [\text{let}(\hole{e};x.\hole{s})]}
            $\begin{array}{c}
             \hat{a} \in \ABT{}_s
            \end{array}$
        \\\\
        \infelong{plus}
            {}
            {[\hat{a}]  \ltrans{\{\text{plus}\}} [\text{plus}(\hole{e};\hole{e})]}
            $\begin{array}{c}
             \hat{a} \in \ABT{}_e 
            \end{array}$
        \\\\
        \infelong{var}
            {}
            {[\hat{a}]  \ltrans{\{\text{var} \ x\}} [x]}
            $\begin{array}{c}
             \hat{a} \in \ABT{}_e
            \end{array}$
    }{\caption{Selected reduction rules for substitution}\label{fig:example_substitution_rules}}
    Notice that in every rule we ensure that the substitution can only be performed if the abt $\hat{a}$ is of the same sort as the operator. For example, given the configuration $\lr{\{let\}.nil, \text{let $x = [\hole{e}]$ in $x+x$}}$ we cannot substitute in a statement, as shown below:
    \begin{center}
    \begin{tabular}{c}
        \inference[(Context)]
            {\inference[(let)]
                {}
                {[\hole{e}]  \ltransn{\{\text{let}\}}}
                $\begin{array}{c}
                    \hole{e} \notin \ABT{}_s
                \end{array}$
            }
            {\lr{\{\text{let}\}.\text{nil}, \text{let $x = [\hole{e}]$ in $x+x$}} \ltransn{\{\text{let}\}}}
    \end{tabular}
    \end{center}

    Similarly, using the general form of cursor movement transition rules in \cref{fig:cursor_movement_rules}, we define specialized rules for every argument $i$ of every operator $o$. In \cref{fig:example_cursor_movement_rules} we show the parent and child rules for the let operator specifically. The transition rules for the remaining operators would be defined analogous to these.
    \federeboks{c}{
        \infelong{letc-1}
            {}
            {[\text{let}(a_1; x.a_2)] \ltrans{\text{child} \ 1} \text{let}([a_1]; x.a_2)}
        \\\\
        \infelong{letc-2}
            {}
            {[\text{let}(a_1; x.a_2)] \ltrans{\text{child} \ 2} \text{let}(a_1; x.[a_2])}
        \\\\
        \infelong{letp-1}
            {}
            {\text{let}([a_1]; x.a_2) \ltrans{\text{parent}} [\text{let}(a_1; x.a_2)]}
        \\\\
        \infelong{letp-2}
            {}
            {\text{let}(a_1; x.[a_2]) \ltrans{\text{parent}} [\text{let}(a_1; x.a_2)]}
    }{\caption{Reduction rules for cursor movement on the $let$ operator}\label{fig:example_cursor_movement_rules}}
\end{example}

\subsubsection{Conditions}
Finally, we define the satisfaction relation for conditions $\phi$ in our editor calculus. The propositional connectives are defined as expected in \cref{fig:satisfaction_relation_connectives}.
\federeboks{c}{
        \infeshort{Negation}
        {[\hat{a}] \nvDash \phi}
        {[\hat{a}] |= \neg \phi}
    \\\\
    \infeshort{Conjunction}
        {[\hat{a}] |= \phi_1 \quad [\hat{a}] |= \phi_2}
        {[\hat{a}] |= \phi_1 \wedge \phi_2}
    \\\\
    \infeshort{Disjunction-1}
        {[\hat{a}] |= \phi_1}
        {[\hat{a}] |= \phi_1 \vee \phi_2}
    \\\\
    \infeshort{Disjunction-2}
        {[\hat{a}] |= \phi_2}
        {[\hat{a}] |= \phi_1 \vee \phi_2}
}{\caption{Satisfaction relation for propositional connectives}\label{fig:satisfaction_relation_connectives}}

The general form of the satisfaction rules for the modalities $@o$, $\Diamond o$ and $\square o$ are defined in \cref{fig:satisfaction_relation_modal}. 
\federeboks{c}{
    \infeshort{At-op}
        {}
        {[o(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_n}.\hat{a}_n)] |= @o }  
    \\\\
    \infeshort{Necessity}
        {[\hat{a}_1] |= \Diamond o \ \dots \ [\hat{a}_n] |= \Diamond o}
        {[o(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_n}.\hat{a}_n)] |=\square o}

    \\\\
    \infeshort{Possibly-i}
        {[\hat{a}_i] |= \Diamond o}
        {[o(\Vec{x_1}.\hat{a}_1; \dots; \Vec{x_i}.\hat{a}_i; \dots; \Vec{x_n}.\hat{a}_n)] |= \Diamond o}
    \\\\
    \infeshort{Possibly-trivial}
        {[\hat{a}] |= @o}
        {[\hat{a}] |= \Diamond o}
}{\caption{Satisfaction relation for modal operators}\label{fig:satisfaction_relation_modal}}
The modal operator $@o$ is satisfied when the root of the abt encapsulated by the cursor is the operator $o$. Therefore we define the rule (At-op) for every operator $o$.

The modal operator $\Diamond o$ is satisfied when the operator $o$ is anywhere within the abt encapsulated by the cursor. For the trivial case, when the root of the tree encapsulated by the cursor is $o$, we only define the one rule (Possibly-trivial). For the non-trivial case, where $o$ is in one of the subtrees $\hat{a}_i$, we define a rule (Possibly-i) for every argument of every operator $o$.

The modal operator $\square o$ is satisfied when $o$ is somewhere in every subtree of the abt encapsulated by the cursor. To describe this, we define the rule (Necessity) for every operator $o$.

\begin{example}\label{ex:modal_logic}
    We finalize our specialized editor calculus from \cref{ex:substitution_rules}, by defining the satisfaction relation. The propositional connectives are defined as previously shown in \cref{fig:satisfaction_relation_connectives}. In \cref{fig:example_satisfaction_relation} we have shown the satisfaction rules for all modalities for the $plus$ operator. Rules for the remaining operators $o$ are defined accordingly.
    \federeboks{c}{
        \infeshort{At-plus}
            {}
            {[\text{plus}(\hat{a}_1;\hat{a}_2)]|= @\text{plus} }  
        \\\\
        \infeshort{Pos1-plus}
            {[\hat{a}_1] |= \Diamond o}
            {[\text{plus}(\hat{a}_1; \hat{a}_2)] |= \Diamond o}
        \\\\
        \infeshort{Nec-plus}
            {[\hat{a}_1] |= \Diamond o \quad [\hat{a}_2] |= \Diamond o}
            {[\text{plus}(\hat{a}_1;\hat{a}_2)] |=\square o}
        \\\\
        \infeshort{Pos2-plus}
            {[\hat{a}_2] |= \Diamond o}
            {[\text{plus}(\hat{a}_1; \hat{a}_2)] |= \Diamond o}
    }{\caption{Satisfaction relation for modal operators on $plus$}\label{fig:example_satisfaction_relation}}
\end{example}

\section{Encoding the generalized editor calculus in an extended lambda calculus} \label{sec:encoding}
In this section we attempt to soundly encode the generalized editor calculus using the simply typed $\lambda$-calculus (\stlc) as the base calculus. The calculus will be extended with various constructs such as pattern matching, pairs and the recursive \textit{fix} operator. To encode the abstract binding trees, we will only need the simply typed lambda calculus. To encode the contexts, we introduce pairs. To encode atomic prefix commands we further extend the calculus with pattern matching ($\lambda_{->,p}$). Lastly to encode modal logic and editor expressions we extend the calculus with a fixed point operator ($\lambda_{->,p,fix}$).

%A concrete overview of the calculus and the constructs necessary to encode every central concept of the editor calculus can be seen in \cref{tab:encoding_requirements}.

%\begin{table}[H]
%    \centering
%    \begin{tabular}{| l | l |}
%        \hline
%        Editor calculus & Encoding\\
%        \hline
%        \ Abstract Binding Trees & \ Simply typed lambda calculus, \stlc \\
%        \hline
%        \ Atomic Prefix Commands & \ \stlc with pattern matching, $\lambda_{->,p}$\\
%        \hline
%        \ Modal Logic & \ \stlc with pattern matching and recursion, $\lambda_{->,p,fix}$ \\
%        \hline
%        \ Editor Expressions & \ \stlc with pattern matching and recursion, $\lambda_{->,p,fix}$ \\
%        \hline 
%    \end{tabular}
%    \vspace{1mm}
%    \caption{Calculus requirements to encode the generalized editor calculus}
%    \label{tab:encoding_requirements}
%\end{table}

\subsection{Motivation}
The motivation behind encoding the generalized editor calculus in $\lambda_{->,p,fix}$ (or subsets thereof) is most importantly the type system it provides. If our encoding is sound, then any instance of the editor calculus will have a sound type system, regardless of the abstract syntax. This follows from the fact that the type system of $\lambda_{->,p,fix}$ is sound. Our editor calculus having a sound type system means that it will reject incorrect editor expressions.

Secondly a $\lambda_{->,p,fix}$ encoding provides a clear strategy for implementing any instantiation of the generalized editor calculus in a functional programming language.

\subsection{Abstract Binding Trees}
To encode abstract binding trees we only need to use the simply typed lambda-calculus. We extend the lambda-calculus with term constants $o$ for every $o \in \OP$ excluding cursors, and the base types $s$ for every $s \in \SORT$. 
Typing rules for operators can be seen on \cref{fig:abt_typing_rules}, where we can infer the type of an operator in the lambda-calculus by its arity. 
\federeboks{C}{
    \infeshort{T-Operator}
        {o \in \OP \text{ and has arity } (\Vec{s_1}.s_1, \dots, \Vec{s_n}.s_n)s}
        {\Gamma |- o : (\Vec{s_1} -> s_1) -> \dots (\Vec{s_n} -> s_n) -> s}
}{
    \caption{Typing rules for operators.}
    \label{fig:abt_typing_rules}
}

With the operators added as term constants and sorts added as base types, encoding abts is straightforward as we just have to curry the operator $o$, encode the children $a_1, \dots, a_n$ and add typings. This can be seen on \cref{fig:enc_abt}.

\federeboks{C}{
    \encode{o(\Vec{x_1}.a_1, \dots , \Vec{x_n}.a_n)} = o \ (\lambda \Vec{x_1} : \Vec{s_1}.\encode{a_1}) \dots (\lambda \Vec{x_n} : \Vec{s_n}.\encode{a_n})
}{
    \caption{Encoding of abstract binding trees.}
    \label{fig:enc_abt}
}

\begin{example}\label{ex:enc_abt_example}
    Continuing from \cref{ex:modal_logic} we will now encode the operators in our small language. Following our singular rule for encoding abt's we get the resulting encoding seen on \cref{fig:enc_abt_example}, and the type for each operator on \cref{fig:abt_example_type}.

    \federeboks{C C}{
        \encode{\text{plus}(a_1; a_2)} = \text{plus} \encode{a_1} \encode{a_2} 
        &
        \encode{\text{num}[n]} = \text{num}[n]
        \\\\
        \encode{\text{var}[x]} = \text{var}[x]
        &
        \encode{\text{hole}_e} = \text{hole}_e
        \\\\
        \encode{\text{exp}(a_1)} = \text{exp} \encode{a_1}
        &
        \encode{\text{hole}_s} = \text{hole}_s
        \\\\
        \multicolumn{2}{C}{
            \encode{\text{let}(a_1; x.a_2)} = \text{let} \ (\encode{a_1}) (\lambda x : e. \encode{a_2})
        }
    }{
        \caption{Encoding of operators.}
        \label{fig:enc_abt_example}
    }

    \federeboks{C C}{
        \text{plus} : e \to e \to e
        &
        \text{num}[n] : e
        \\\\
        \text{var}[x] : e
        &
        \text{hole}_e : e
        \\\\
        \text{exp} : e \to s
        &
        \text{hole}_s : s
        \\\\
        \multicolumn{2}{C}{
            \text{let} : e \to (e \to s) \to s
        }
    }{
        \caption{Types for the encoded operators.}
        \label{fig:abt_example_type}
    }

\end{example}

for example, if we were to encode the statement:
\begin{equation*}
        \text{let $x=[\hole{e}]$ in $x+5$}
    \end{equation*}
    it could be encoded as:
    \begin{align*} 
        \text{let} \ (\text{cursor}_e(\text{hole}_e)) \ (\lambda x . \text{plus} \ \text{var}[x] \ \text{num}[5])
  \end{align*}

\subsection{Cursor Contexts}\label{sec:enc_cursor_contexts}

To represent abts with cursor contexts we introduce the pair construct similarly to \cite{types_programming_languages}.
%In \cref{fig:abt_cursor_syntactic_forms} we introduce the new syntactic forms which include the pair construct, first projection, second projection and the product type. 
%\federeboks{R C L l}{
%    \multicolumn{4}{c}{Terms}
%    \\
%    M & ::= &(M_1,M_2) & (pair) \\
%      & |   & M.1      & (first projection) \\
%      & |   & M.2      & (second projection)
%    \\\\
%    \multicolumn{4}{c}{Types}
%    \\
%    \tau & ::= & \tau_1 \times \tau_2 & (product type)
%}{\caption{Abstract syntax for representing cursor contexts as pairs.}\label{fig:abt_cursor_syntactic_forms}}

In \cref{fig:enc_abt_cursor} the encoding for cursor contexts can be seen. We can encode a cursor context $C[a]$ as a pair of two encoded trees the first being a regular abt and the second being an abt with a context hole $\ctxhole$, which we encode as an operator $\ctxhole$. We also introduce a type alias such that $\encode{C[a]}$ has type $Ctx = s \times s$.
\federeboks{C}{
    \encode{C[a]} = (\encode{a}, \encode{C}) \\
    \encode{\ctxhole} = \ctxhole
}{
    \caption{Encoding of cursor contexts.}
    \label{fig:enc_abt_cursor}
}

%\begin{example}
%    In \cref{ex:well_formed_trees} we saw that the expression.
%    \begin{equation*}
%        \text{let $x=[\hole{e}]$ in $x+5$}
%    \end{equation*}
%    is well-formed since it can be written as $C[\dot{e}]$ where
 %   \begin{align*}
 %       C &= \quad \text{let $x=\ctxhole$ in $x+5$} \\
  %      \dot{e}&= \quad [\hole{e}]
  %  \end{align*}
%    Encoding $C[\dot{e}]$ following the rules for encoding contexts and the rules for encoding abts in \cref{ex:enc_abt_example} yields the following encoding:
%    \begin{align*}
 %       & \encode{C[\dot{e}]} = \\ 
 %       & (\encode{\text{cursor}_e(\text{hole}_e)}, \encode{\text{let}([\cdot]; x.\text{plus}(\text{var}[x], \text{num}[5]))}) = \\ 
 %       & (\text{cursor}_e(\text{hole}_e), \text{let} \ [\cdot] \ (\lambda x . %\text{plus} \ \text{var}[x] \ \text{num}[5]))
 %   \end{align*}

%\end{example}

\subsection{Atomic Prefix Commands}
To encode the atomic prefix commands we further extend the lambda calculus with pattern matching. As seen in \cref{fig:apc_syntactic_forms} we introduce the match construct where we match the term $M$ on the patterns $p$ resulting in the corresponding term $N$ if there is a match. As mentioned $p$ is the syntactic category describing the pattern we are trying to match. These patterns consist of variables, wildcards, operators, pairs, and lastly bindings which recursively match on the body of an abstraction. 
\federeboks{R C L l}{
    \multicolumn{4}{c}{Terms}
    \\
    M, N & ::= & \text{match}\ M\ \overrightarrow{p \to N} & (match construct)
    \\
    p & ::= & x              & (variable) \\
      & |   & \_             & (wildcard) \\
      & |   & o\ \Vec{p}     & (operator) \\
      & |   & (p_1, p_2)     & (pair) \\
      & |   & .p             & (binding)
}{\caption{Abstract syntax for representing atomic prefix commands, extending the lambda-calculus with a match construct.}\label{fig:apc_syntactic_forms}}
The reduction rules for the match construct can be seen in the full version of the paper.

We now define the auxiliary functions as seen in \cref{tab:zip_oper}. These function as the operations of a zipper data structure, which helps encode atomic prefix commands on abts. We also introduce the abbreviation $[a]$ to mean the cursor operator with the child $a$.

\federeboks{L}{
    $down$ \define \lambda x:s.\text{match }x
    \quad [o \ (.a_1) \dots (.a_n)] \to \\
    \quad o \ (\lambda \Vec{x_1}.[a_1]) \dots (\Vec{x_n}.a_n) \\\\

    $right$ \define \lambda x:s.\text{match }x
    \quad o \ (.a_1) \ \dots (.[a_i]) \ \dots (.a_n) \to \\
    \quad o \ (\lambda \Vec{x_1}.a_1) \ \dots (\lambda \Vec{x_{i+1}}.[a_{i+1}]) \ \dots (\Vec{x_n}.a_n) \\\\

    $up$ \define \lambda x:s.\text{match }x
    \quad o \ (.a_1) \dots (.[a_i]) \dots (.a_n) \to \\
    \quad [o \ (\lambda \Vec{x_1}.a_1) \dots (\lambda \Vec{x_n}.a_n)] \\\\

    $set$ \define \lambda a:s.\lambda x:s.\text{match }x
    \quad [a'] \to [a]
}{\caption{Function definitions for cursor movement and substitution on abstract binding trees.}\label{tab:zip_oper}}
The encoding of atomic prefix commands can be seen in \cref{fig:atc_enc}. The $child$ command can be encoded as a base case and a recursive case. In the base case we want to move to the first $child$, which means we just go down. In the recursive case we want to move to $child$ $n$, we can encode this as moving to $child$ $n-1$ and then moving right. $parent$ and $insert$ can directly be encoded as $up$ and $set$ respectively.
\federeboks{L L}{
    \encode{\text{child} \ 1} = \text{down}
    &
    \encode{\text{child} \ n} = \text{right}\ \encode{\text{child} \ n - 1}
    \\\\
    \encode{\text{parent}} = \text{up}
    &
    \encode{\text{insert} \ a} = \text{set} \ \encode{a}
}{
    \caption{Encoding of atomic prefix commands.}
    \label{fig:atc_enc}
}

\subsection{Control Structures and Modal Logic}
To encode control structures and modal logic, we need to further extend the lambda-calculus with a \fix operator as well as the $Bool$ base type. We introduce the \fix operator as seen in \cite{types_programming_languages}, as well as boolean types and values. 
On \cref{fig:edt_enc} the encoding of editor expressions can be seen. For atomic prefixes the atomic prefix command should be applied before the rest of the editor expression. The $nil$ editor expression will become the identity function. Sequential editor expressions will apply the encoding of the editor expressions in the correct order. Recursion works by using the newly introduced \fix operator. Lastly, conditional editor expressions are encoded such that we match on the evaluation of $\phi$ on the first projection of a context $C$ and apply the encoding of $E_1$ or $E_2$ on $C$ accordingly. For the full encoding of modal logic, we refer to the full version of the paper.
%\federeboks{R C L l}{
%    \multicolumn{4}{c}{Terms}
%    \\
%    M & ::= & \fix\ M & (fixed point of M) \\
%      & |   & \top    & (true) \\
%      & |   & \bot    & (false)
%    \\\\
%    p & ::= & \top & (match true) \\
%      & |   & \bot & (match false)
%    \\\\
%    \multicolumn{4}{c}{Types}
%    \\
%    \tau & ::= & \text{Bool} & (boolean)
%}{\caption{Abstract syntax for control structures and modal logic, extending the calculus with the \fix operator and boolean types and term constants.}\label{fig:control_syntactic_forms}}

%In \cref{fig:cond_aux} we define auxiliary functions that will help with the encoding of conditional expressions. The auxiliary functions $and$, $or$ and $neg$ function as their name would suggest. In our semantics, when checking if a formula $\phi$ holds, we do so against an abt $a$ (i.e $a \vDash \phi$). When constructing the proof tree for such a formula we see that this abt $a$ is propagated up through the proof tree, thus we need to have the ability to propagate an abt throughout our encoded formula as well. To propagate this abt through $and$, $or$ and $neg$ we define the combinators $phoenix$ (liftA2/liftM2 in Haskell) and $bluebird$ (function composition / B combinator).

\federeboks{C C}{
    \encode{\pi.E} = \lambda C:\text{Ctx}.\encode{E} ( (\encode{\pi} \ C.1) \ , C.2) 
    \\\\
    \encode{E_1 \ggg E_2}  = \lambda C:\text{Ctx}. \encode{E_2} (\encode{E_1} C)
    \\\\
    \encode{\text{rec } x.E} = \text{fix } ( \lambda x: (\text{Ctx} \to \text{Ctx}).\encode{E}) 
    \\\\
    \encode{\phi => E_1 | E_2} = \lambda C:\text{Ctx}.\text{match} \ (\encode{\phi} \ C.1) \\
        | \ \top ->  \encode{E_1} \ C 
    \\
    | \  \bot ->  \encode{E_2} \ C 
    \\\\
    \encode{\lr{E, C[a']}} = \encode{E} \  (\encode{a'}, \encode{C})
        \\\\
    \encode{\text{nil}} = \lambda C:\text{Ctx}.C \quad \quad    \encode{x} = x

}{
    \caption{Encoding of editor expressions and context configuration.}
    \label{fig:edt_enc}
}

\iffalse
\federeboks{L C L}{
    \text{phoenix} &\define& \lambda f: (\text{Bool} \to \text{Bool} \to \text{Bool}). \\
            &       & \lambda g: (s \to \text{Bool}). \\
            &       & \lambda h: (s \to \text{Bool}) . \\
            &       & \lambda a:s . \\
            &       & f (g a) (h a) 
    \\\\
    \text{bluebird} &\define& \lambda f: (\text{Bool} \to \text{Bool}). \\
             &       & \lambda g: (s \to \text{Bool}) . \\
             &       & a:s . \\
             &       & f (g a)
    \\\\

    \iffalse
    \text{and} &\define& \lambda \phi_1:(s -> \text{Bool}). \lambda \phi_2:(s -> \text{Bool}) .\lambda a:s.\text{match} (\phi_1 \ a, \phi_2 \ a) \\&&(\top, \top) -> \top \\&|& (\_,\_) -> \bot 
    \\\\

    \text{or} &\define& \lambda \phi_1:(s -> \text{Bool}). \lambda \phi_2:(s -> \text{Bool}) .\lambda a:s.\text{match} (\phi_1 \ a, \phi_2 \ a)\\&& (\bot, \bot) -> \bot \\&|& (\_,\_) -> \top
    \\\\

    \text{neg} &\define& \lambda \phi:(s -> \text{Bool}).\lambda a:s.\text{match} (\phi \ a) \\
    &|& \top -> \bot \\
    &|& \bot -> \top
    \\\\
    \fi

    \text{or} &\define& \lambda b_1:\text{Bool}. \lambda b_2:\text{Bool}.\text{match} \  (b1,b2) \ \\
    &|& (\bot,\bot) -> \bot \\
    &|& (\_, \_) -> \top)
    \\\\

    \text{and} &\define& \lambda b_1:\text{Bool}. \lambda b_2:\text{Bool}.\text{match} \  (b1,b2) \\
    &|& (\top,\top) -> \top \\
    &|& (\_, \_) -> \bot)
    \\\\

    \text{neg} &\define& \lambda b:\text{Bool} . \text{match } b \\
        & |     & \top \to \bot \\
        & |     & \bot \to \top
}{
    \caption{Function definitions for auxiliary functions.}
    \label{fig:cond_aux}
}
The encoding of conjunction, disjunction and negation, shown in \cref{fig:encoding_connectives}, comes quite naturally using the previously defined auxiliary functions.
\federeboks{L}{
    \encode{\phi_1 \wedge \phi_2 } = \text{phoenix and} \ \encode{\phi_1} \ \encode{\phi_2}
    \\\\

    \encode{\phi_1 \vee \phi_2} = \text{phoenix or} \ \encode{\phi_1} \ \encode{\phi_2}
    \\\\

    \encode{\neg \phi} = \text{bluebird neg} \ \encode{\phi}
    \\\\
    \encode{@o} = \lambda x:s . \text{match } x \\
                \quad | \ [o \ \_ \ \dots \ \_] \to \top  \\
                \quad | \ \_ \to \bot
}{\caption{Encoding of propositional connectives}\label{fig:encoding_connectives}}
In \cref{fig:cond_enc} we show the encoding of modal logic operators.
\federeboks{L}{    
    \encode{\Diamond o} = \text{\fix} (\lambda f:(s->\text{Bool})
    . \lambda x:s . \text{match } x \\
                        \quad | \ [o \ \_ \ \dots \ \_] \to \top \\
                        \quad | \  [\_(\_.a_1 , \dots ,\_.a_n)] \to \\
                        \quad\ \ \text{or} ( \dots (\text{or}\ (f\ [a_1]) \ (f \ [a_2]))\ \dots ) \ (f\ [a_n])\\
                        \quad | \ \_ \to \bot )
    \\\\
    \encode{\square o_1} = \lambda x:s . \text{match } x \\
                         \quad | \ \_(\_.a_1 , \dots ,\_.a_n) \to \\
                         \quad\ \ \text{and}(\dots (\text{and}  (\encode{\Diamond o}\ a_1) (\encode{\Diamond o} \ a_2)) \dots )\ (\encode{\Diamond o} \ a_n) \\
                         \quad | \ \_ \to \top
}{
    \caption{Encoding of modal logic and boolean logic operators.}
    \label{fig:cond_enc}
}
$@o$ is encoded such that we pattern match on the on the operator $o$, and if there is a match the resulting value is true, otherwise it is false. $\Diamond o$ uses the \fix operator, as the function is recursively defined. 
$\square o$ consists of either applying $\Diamond o$ on all the children of the tree, or returning true if there are no children.
\fi
With this, the entire editor calculus has been encoded. We refer to the full version of the paper to see the proof of the soundness of our encoding.
\iffalse
\begin{example}
    To illustrate how editor expressions are encoded, we provide an encoding of the following expression
    \begin{equation*}
        \text{child\ $2$.nil $\ggg$ parent.nil}
    \end{equation*}
    using the encoding rules seen on \cref{fig:edt_enc} and \cref{fig:atc_enc}. For readability purposes we will not add explicit typings, just note that any variable $C_i$ has type $Ctx$.
    \begin{align*}
         &\ \encode{\text{child\ $2$.nil $\ggg$ parent.nil}} \\
         =&\ \lambda C_1 . \encode{\text{parent.nil}} (\encode{\text{child\ $2$.nil}} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . \encode{\text{nil}}(\encode{\text{parent}} C_2.1, C_2.2)) (\encode{\text{child\ $2$.nil}} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . \encode{nil}(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (\lambda C_3 . C_3)(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
         \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . \encode{nil} (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\lambda C_4.C_4) (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
         \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (\encode{child\ 1} (C_3.1, C_3.2))C_1) \\
         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (down (C_3.1, C_3.2))C_1) \\
    \end{align*}
    As can be seen, the editor expression could be fully encoded to the extended lambda calculus. We refer to the full paper for
\end{example}
\fi

\section{Conclusion \& Further Work}
%Conclusion
We have developed a generalized editor calculus that enables the creation of a syntax-directed editor calculus for a specific abstract syntax. Subsequently, we encoded this editor calculus into an extended version of the lambda calculus, which incorporates pairs, recursion and pattern matching.
Further work on our generalized editor calculus could be to prove whether the encoding is complete. Otherwise, one could implement a redo-undo construct such as introduced in \cite{undo_Edit} in the generalized editor calculus. Another interesting construct to implement would be a copy-paste operation.

%Furthermore, we have proven the soundness of the encoding of the generalized editor calculus into the extended lambda calculus. As a consequence, given that the type system of the extended lambda calculus is sound, it logically follows that our editor calculus also possesses a sound type system.

%We have identified several possibilities for expanding our work on the generalized editor calculus. First of all, it would be excellent to prove that our encoding is complete. However, in the scope of this paper, we decided to only include the proof of the soundness of the encoding, as that would give us the sound type system.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2023-06-30 sample-guide.tex
Filerne er tilføjet nu

%%
%% This is file `sample-sigplan.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigplan')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigplan.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%%
%% Commands for TeXCount
%TC:macro \cite [option:text,text]
%TC:macro \citep [option:text,text]
%TC:macro \citet [option:text,text]
%TC:envir table 0 1
%TC:envir table* 0 1
%TC:envir tabular [ignore] word
%TC:envir displaymath 0 word
%TC:envir math 0 word
%TC:envir comment 0 0
%%
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
%% For submission and review of your manuscript please change the
%% command to \documentclass[manuscript, screen, review]{acmart}.
%%
%% When submitting camera ready or to TAPS, please change the command
%% to \documentclass[sigconf]{acmart} or whichever template is required
%% for your publication.
%%
%%
\documentclass[sigplan,screen]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{XXXXXXX.XXXXXXX}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Conference acronym 'XX]{Make sure to enter the correct
  conference title from your rights confirmation emai}{June 03--05,
  2018}{Woodstock, NY}
%%
%%  Uncomment \acmBooktitle if the title of the proceedings is different
%%  from ``Proceedings of ...''!
%%
%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
%%  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{A generalised editor calculus}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

%%\orcid{1234-5678-9012}

\author{Andreas Tor Mortensen}
\email{atmo20@student.aau.dk}

\author{Benjamin Bennetzen}
\email{bbenne20@student.aau.dk}

\author{Nikolaj Rossander Kristensen}
\email{nrkr20@student.aau.dk}

\author{Peter Buus Steffensen}
%%\authornotemark[1]
\email{psteff19@student.aau.dk
}

\affiliation{%
  \institution{Department of Computer Science at Aalborg University}
  \streetaddress{Selma Lagerlöfsvej 300}
  \city{Aalborg}
  \country{Denmark}
  \postcode{9220}
}

\author{Hans Huttel}
\email{hans@cs.aau.dk}
\affiliation{%
  \institution{Department of Computer Science at Aalborg University}
  \streetaddress{Selma Lagerlöfsvej 300}
  \city{Aalborg}
  \country{Denmark}
  \postcode{9220}
}

%\author{Lars Th{\o}rv{\"a}ld}
%\affiliation{%
 % \institution{The Th{\o}rv{\"a}ld Group}
%  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
%  \city{Hekla}
%  \country{Iceland}}
%\email{larst@affiliation.org}

%\author{Valerie B\'eranger}
%\affiliation{%
%  \institution{Inria Paris-Rocquencourt}
%  \city{Rocquencourt}
%  \country{France}
%}

%\author{Aparna Patel}
%\affiliation{%
% \institution{Rajiv Gandhi University}
% \streetaddress{Rono-Hills}
 %\city{Doimukh}
 %\state{Arunachal Pradesh}
% \country{India}}

%\author{Huifen Chan}
%\affiliation{%
  %\institution{Tsinghua University}
  %\streetaddress{30 Shuangqing Rd}
 % \city{Haidian Qu}
%  \state{Beijing Shi}
%  \country{China}}

%\author{Charles Palmer}
%\affiliation{%
 % \institution{Palmer Research Laboratories}
%  \streetaddress{8600 Datapoint Drive}
 % \city{San Antonio}
 % \state{Texas}
 % \country{USA}
  %\postcode{78229}}
%\email{cpalmer@prl.com}

%\author{John Smith}
%\affiliation{%
 % \institution{The Th{\o}rv{\"a}ld Group}
  %\streetaddress{1 Th{\o}rv{\"a}ld Circle}
 % \city{Hekla}
 % \country{Iceland}}
%\email{jsmith@affiliation.org}

%\author{Julius P. Kumquat}
%\affiliation{%
 % \institution{The Kumquat Consortium}
%  \city{New York}
 % \country{USA}}
%\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
In this paper, we present a generalization of a syntax-directed editor calculus, which can be used to instantiate a specialized syntax-directed editor for any language with explicit binders, given by some abstract syntax. The editor calculus guarantees the absence of syntactical errors while allowing incomplete programs. The generalized editor calculus is then encoded into a simply typed lambda calculus, extended with pairs, booleans, pattern matching and fixed points.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Syntax-directed editor, Generalisation, lambda calculus}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
\begin{teaserfigure}
  \includegraphics[width=\textwidth]{Images/windmill.jpg}
  \caption{Seattle Mariners at Spring Training, 2010.}
  \Description{Enjoying the baseball game from the third-base
  seats. Ichiro Suzuki preparing to bat.}
  \label{fig:teaser}
\end{teaserfigure}

\received{20 February 2007}
\received[revised]{12 March 2009}
\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
ACM's consolidated article template, introduced in 2017, provides a
consistent \LaTeX\ style for use across ACM publications, and
incorporates accessibility and metadata-extraction functionality
necessary for future Digital Library endeavors. Numerous ACM and
SIG-specific \LaTeX\ templates have been examined, and their unique
features incorporated into this single new template.

If you are new to publishing with ACM, this document is a valuable
guide to the process of preparing your work for publication. If you
have published with ACM before, this document provides insight and
instruction into more recent changes to the article template.

The ``\verb|acmart|'' document class can be used to prepare articles
for any ACM publication --- conference or journal, and for any stage
of publication, from review to final ``camera-ready'' copy, to the
author's own version, with {\itshape very} few changes to the source.

\section{Template Overview}
As noted in the introduction, the ``\verb|acmart|'' document class can
be used to prepare many different kinds of documentation --- a
double-blind initial submission of a full-length technical paper, a
two-page SIGGRAPH Emerging Technologies abstract, a ``camera-ready''
journal article, a SIGCHI Extended Abstract, and more --- all by
selecting the appropriate {\itshape template style} and {\itshape
  template parameters}.

This document will explain the major features of the document
class. For further information, the {\itshape \LaTeX\ User's Guide} is
available from
\url{https://www.acm.org/publications/proceedings-template}.

\subsection{Template Styles}

The primary parameter given to the ``\verb|acmart|'' document class is
the {\itshape template style} which corresponds to the kind of publication
or SIG publishing the work. This parameter is enclosed in square
brackets and is a part of the {\verb|documentclass|} command:
\begin{verbatim}
  \documentclass[STYLE]{acmart}
\end{verbatim}

Journals use one of three template styles. All but three ACM journals
use the {\verb|acmsmall|} template style:
\begin{itemize}
\item {\texttt{acmsmall}}: The default journal template style.
\item {\texttt{acmlarge}}: Used by JOCCH and TAP.
\item {\texttt{acmtog}}: Used by TOG.
\end{itemize}

The majority of conference proceedings documentation will use the {\verb|acmconf|} template style.
\begin{itemize}
\item {\texttt{acmconf}}: The default proceedings template style.
\item{\texttt{sigchi}}: Used for SIGCHI conference articles.
\item{\texttt{sigplan}}: Used for SIGPLAN conference articles.
\end{itemize}

\subsection{Template Parameters}

In addition to specifying the {\itshape template style} to be used in
formatting your work, there are a number of {\itshape template parameters}
which modify some part of the applied template style. A complete list
of these parameters can be found in the {\itshape \LaTeX\ User's Guide.}

Frequently-used parameters, or combinations of parameters, include:
\begin{itemize}
\item {\texttt{anonymous,review}}: Suitable for a ``double-blind''
  conference submission. Anonymizes the work and includes line
  numbers. Use with the \texttt{\acmSubmissionID} command to print the
  submission's unique ID on each page of the work.
\item{\texttt{authorversion}}: Produces a version of the work suitable
  for posting by the author.
\item{\texttt{screen}}: Produces colored hyperlinks.
\end{itemize}

This document uses the following string as the first command in the
source file:
\begin{verbatim}
\documentclass[sigplan,screen]{acmart}
\end{verbatim}

\section{Modifications}

Modifying the template --- including but not limited to: adjusting
margins, typeface sizes, line spacing, paragraph and list definitions,
and the use of the \verb|\vspace| command to manually adjust the
vertical spacing between elements of your work --- is not allowed.

{\bfseries Your document will be returned to you for revision if
  modifications are discovered.}

\section{Typefaces}

The ``\verb|acmart|'' document class requires the use of the
``Libertine'' typeface family. Your \TeX\ installation should include
this set of packages. Please do not substitute other typefaces. The
``\verb|lmodern|'' and ``\verb|ltimes|'' packages should not be used,
as they will override the built-in typeface families.

\section{Title Information}

The title of your work should use capital letters appropriately -
\url{https://capitalizemytitle.com/} has useful rules for
capitalization. Use the {\verb|title|} command to define the title of
your work. If your work has a subtitle, define it with the
{\verb|subtitle|} command.  Do not insert line breaks in your title.

If your title is lengthy, you must define a short version to be used
in the page headers, to prevent overlapping text. The \verb|title|
command has a ``short title'' parameter:
\begin{verbatim}
  \title[short title]{full title}
\end{verbatim}

\section{Authors and Affiliations}

Each author must be defined separately for accurate metadata
identification.  As an exception, multiple authors may share one
affiliation. Authors' names should not be abbreviated; use full first
names wherever possible. Include authors' e-mail addresses whenever
possible.

Grouping authors' names or e-mail addresses, or providing an ``e-mail
alias,'' as shown below, is not acceptable:
\begin{verbatim}
  \author{Brooke Aster, David Mehldau}
  \email{dave,judy,steve@university.edu}
  \email{firstname.lastname@phillips.org}
\end{verbatim}

The \verb|authornote| and \verb|authornotemark| commands allow a note
to apply to multiple authors --- for example, if the first two authors
of an article contributed equally to the work.

If your author list is lengthy, you must define a shortened version of
the list of authors to be used in the page headers, to prevent
overlapping text. The following command should be placed just after
the last \verb|\author{}| definition:
\begin{verbatim}
  \renewcommand{\shortauthors}{McCartney, et al.}
\end{verbatim}
Omitting this command will force the use of a concatenated list of all
of the authors' names, which may result in overlapping text in the
page headers.

The article template's documentation, available at
\url{https://www.acm.org/publications/proceedings-template}, has a
complete explanation of these commands and tips for their effective
use.

Note that authors' addresses are mandatory for journal articles.

\section{Rights Information}

Authors of any work published by ACM will need to complete a rights
form. Depending on the kind of work, and the rights management choice
made by the author, this may be copyright transfer, permission,
license, or an OA (open access) agreement.

Regardless of the rights management choice, the author will receive a
copy of the completed rights form once it has been submitted. This
form contains \LaTeX\ commands that must be copied into the source
document. When the document source is compiled, these commands and
their parameters add formatted text to several areas of the final
document:
\begin{itemize}
\item the ``ACM Reference Format'' text on the first page.
\item the ``rights management'' text on the first page.
\item the conference information in the page header(s).
\end{itemize}

Rights information is unique to the work; if you are preparing several
works for an event, make sure to use the correct set of commands with
each of the works.

The ACM Reference Format text is required for all articles over one
page in length, and is optional for one-page articles (abstracts).

\section{CCS Concepts and User-Defined Keywords}

Two elements of the ``acmart'' document class provide powerful
taxonomic tools for you to help readers find your work in an online
search.

The ACM Computing Classification System ---
\url{https://www.acm.org/publications/class-2012} --- is a set of
classifiers and concepts that describe the computing
discipline. Authors can select entries from this classification
system, via \url{https://dl.acm.org/ccs/ccs.cfm}, and generate the
commands to be included in the \LaTeX\ source.

User-defined keywords are a comma-separated list of words and phrases
of the authors' choosing, providing a more flexible way of describing
the research being presented.

CCS concepts and user-defined keywords are required for for all
articles over two pages in length, and are optional for one- and
two-page articles (or abstracts).

\section{Sectioning Commands}

Your work should use standard \LaTeX\ sectioning commands:
\verb|section|, \verb|subsection|, \verb|subsubsection|, and
\verb|paragraph|. They should be numbered; do not remove the numbering
from the commands.

Simulating a sectioning command by setting the first word or words of
a paragraph in boldface or italicized text is {\bfseries not allowed.}

\section{Tables}

The ``\verb|acmart|'' document class includes the ``\verb|booktabs|''
package --- \url{https://ctan.org/pkg/booktabs} --- for preparing
high-quality tables.

Table captions are placed {\itshape above} the table.

Because tables cannot be split across pages, the best placement for
them is typically the top of the page nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and the
table caption.  The contents of the table itself must go in the
\textbf{tabular} environment, to be aligned properly in rows and
columns, with the desired horizontal and vertical rules.  Again,
detailed instructions on \textbf{tabular} material are found in the
\textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed output of
this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more
desirable. Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}

Always use midrule to separate table header rows from data rows, and
use it only for this purpose. This enables assistive technologies to
recognise table headers and support their users in navigating tables
more easily.

\section{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of the three are
discussed in the next sections.

\subsection{Inline (In-text) Equations}
A formula that appears in the running text is called an inline or
in-text formula.  It is produced by the \textbf{math} environment,
which can be invoked with the usual
\texttt{{\char'134}begin\,\ldots{\char'134}end} construction or with
the short form \texttt{\$\,\ldots\$}. You can use any of the symbols
and structures, from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a few
examples of in-text equations in context. Notice how this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols and
structures available in \LaTeX\@; this section will just give a couple
of examples of display equations in context.  First, consider the
equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\section{Figures}

The ``\verb|figure|'' environment should be used for figures. One or
more images can be placed within a figure. If your figure contains
third-party material, you must clearly identify it as such, as shown
in the example below.

Your figures should contain a caption which describes the figure to
the reader.

Figure captions are placed {\itshape below} the figure.

Every figure should also have a figure description unless it is purely
decorative. These descriptions convey what’s in the image to someone
who cannot see it. They are also used by search engine crawlers for
indexing images, and when images cannot be loaded.

A figure description must be unformatted plain text less than 2000
characters long (including spaces).  {\bfseries Figure descriptions
  should not repeat the figure caption – their purpose is to capture
  important information that is not already provided in the caption or
  the main text of the paper.} For figures that convey important and
complex new information, a short text description may not be
adequate. More complex alternative descriptions can be placed in an
appendix and referenced in a short figure description. For example,
provide a data table capturing the information in a bar chart, or a
structured list representing a graph.  For additional information
regarding how best to write figure descriptions and why doing this is
so important, please see
\url{https://www.acm.org/publications/taps/describing-figures/}.

\subsection{The ``Teaser Figure''}

A ``teaser figure'' is an image, or set of images in one figure, that
are placed after all author and affiliation information, and before
the body of the article, spanning the page. If you wish to have such a
figure in your article, place the command immediately before the
\verb|\maketitle| command:
\begin{verbatim}
  \begin{teaserfigure}
    \includegraphics[width=\textwidth]{sampleteaser}
    \caption{figure caption}
    \Description{figure description}
  \end{teaserfigure}
\end{verbatim}

\section{Citations and Bibliographies}

The use of \BibTeX\ for the preparation and formatting of one's
references is strongly recommended. Authors' names should be complete
--- use full first names (``Donald E. Knuth'') not initials
(``D. E. Knuth'') --- and the salient identifying features of a
reference should be included: title, year, volume, number, pages,
article DOI, etc.

The bibliography is included in your source document with these two
commands, placed just before the \verb|\end{document}| command:
\begin{verbatim}
  \bibliographystyle{ACM-Reference-Format}
  \bibliography{bibfile}
\end{verbatim}
where ``\verb|bibfile|'' is the name, without the ``\verb|.bib|''
suffix, of the \BibTeX\ file.

Citations and references are numbered by default. A small number of
ACM publications have citations and references formatted in the
``author year'' style; for these exceptions, please include this
command in the {\bfseries preamble} (before the command
``\verb|\begin{document}|'') of your \LaTeX\ source:
\begin{verbatim}
  \citestyle{acmauthoryear}
\end{verbatim}

  Some examples.  A paginated journal article \cite{Abril07}, an
  enumerated journal article \cite{Cohen07}, a reference to an entire
  issue \cite{JCohen96}, a monograph (whole book) \cite{Kosiur01}, a
  monograph/whole book in a series (see 2a in spec. document)
  \cite{Harel79}, a divisible-book such as an anthology or compilation
  \cite{Editor00} followed by the same example, however we only output
  the series if the volume number is given \cite{Editor00a} (so
  Editor00a's series should NOT be present since it has no vol. no.),
  a chapter in a divisible book \cite{Spector90}, a chapter in a
  divisible book in a series \cite{Douglass98}, a multi-volume work as
  book \cite{Knuth97}, a couple of articles in a proceedings (of a
  conference, symposium, workshop for example) (paginated proceedings
  article) \cite{Andler79, Hagerup1993}, a proceedings article with
  all possible elements \cite{Smith10}, an example of an enumerated
  proceedings article \cite{VanGundy07}, an informally published work
  \cite{Harel78}, a couple of preprints \cite{Bornmann2019,
    AnzarootPBM14}, a doctoral dissertation \cite{Clarkson85}, a
  master's thesis: \cite{anisi03}, an online document / world wide web
  resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game
  (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03} and \cite{Lee05}
  and (Case 3) a patent \cite{JoeScientist001}, work accepted for
  publication \cite{rous08}, 'YYYYb'-test for prolific author
  \cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might
  contain 'duplicate' DOI and URLs (some SIAM articles)
  \cite{Kirschmer:2010:AEI:1958016.1958018}. Boris / Barbara Beeton:
  multi-volume works as books \cite{MR781536} and \cite{MR781537}. A
  couple of citations with DOIs:
  \cite{2004:ITE:1009386.1010128,Kirschmer:2010:AEI:1958016.1958018}. Online
  citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.
  Artifacts: \cite{R} and \cite{UMassCitations}.

\section{Acknowledgments}

Identification of funding sources and other support, and thanks to
individuals and groups that assisted in the research and the
preparation of the work should be included in an acknowledgment
section, which is placed just before the reference section in your
document.

This section has a special environment:
\begin{verbatim}
  \begin{acks}
  ...
  \end{acks}
\end{verbatim}
so that the information contained therein can be more easily collected
during the article metadata extraction phase, and to ensure
consistency in the spelling of the section heading.

Authors should not prepare this section as a numbered or unnumbered {\verb|\section|}; please use the ``{\verb|acks|}'' environment.

\section{Appendices}

If your work needs an appendix, add it before the
``\verb|\end{document}|'' command at the conclusion of your source
document.

Start the appendix with the ``\verb|appendix|'' command:
\begin{verbatim}
  \appendix
\end{verbatim}
and note that in the appendix, sections are lettered, not
numbered. This document has two appendices, demonstrating the section
and subsection identification method.

\section{Multi-language papers}

Papers may be written in languages other than English or include
titles, subtitles, keywords and abstracts in different languages (as a
rule, a paper in a language other than English should include an
English title and an English abstract).  Use \verb|language=...| for
every language used in the paper.  The last language indicated is the
main language of the paper.  For example, a French paper with
additional titles and abstracts in English and German may start with
the following command
\begin{verbatim}
\documentclass[sigconf, language=english, language=german,
               language=french]{acmart}
\end{verbatim}

The title, subtitle, keywords and abstract will be typeset in the main
language of the paper.  The commands \verb|\translatedXXX|, \verb|XXX|
begin title, subtitle and keywords, can be used to set these elements
in the other languages.  The environment \verb|translatedabstract| is
used to set the translation of the abstract.  These commands and
environment have a mandatory first argument: the language of the
second argument.  See \verb|sample-sigconf-i13n.tex| file for examples
of their usage.

\section{SIGCHI Extended Abstracts}

The ``\verb|sigchi-a|'' template style (available only in \LaTeX\ and
not in Word) produces a landscape-orientation formatted article, with
a wide left margin. Three environments are available for use with the
``\verb|sigchi-a|'' template style, and produce formatted output in
the margin:
\begin{description}
\item[\texttt{sidebar}:]  Place formatted text in the margin.
\item[\texttt{marginfigure}:] Place a figure in the margin.
\item[\texttt{margintable}:] Place a table in the margin.
\end{description}

%%
%% The acknowledgments section is defined using the "acks" environment
%% (and NOT an unnumbered section). This ensures the proper
%% identification of the section in the article metadata, and the
%% consistent spelling of the heading.
\begin{acks}
To Robert, for the bagels and explaining CMYK and color spaces.
\end{acks}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-base}

%%
%% If your work has an appendix, this is the place to put it.
\appendix

\section{Research Methods}

\subsection{Part One}

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Morbi
malesuada, quam in pulvinar varius, metus nunc fermentum urna, id
sollicitudin purus odio sit amet enim. Aliquam ullamcorper eu ipsum
vel mollis. Curabitur quis dictum nisl. Phasellus vel semper risus, et
lacinia dolor. Integer ultricies commodo sem nec semper.

\subsection{Part Two}

Etiam commodo feugiat nisl pulvinar pellentesque. Etiam auctor sodales
ligula, non varius nibh pulvinar semper. Suspendisse nec lectus non
ipsum convallis congue hendrerit vitae sapien. Donec at laoreet
eros. Vivamus non purus placerat, scelerisque diam eu, cursus
ante. Etiam aliquam tortor auctor efficitur mattis.

\section{Online Resources}

Nam id fermentum dui. Suspendisse sagittis tortor a nulla mollis, in
pulvinar ex pretium. Sed interdum orci quis metus euismod, et sagittis
enim maximus. Vestibulum gravida massa ut felis suscipit
congue. Quisque mattis elit a risus ultrices commodo venenatis eget
dui. Etiam sagittis eleifend elementum.

Nam interdum magna at lectus dignissim, ac dignissim lorem
rhoncus. Maecenas eu arcu ac neque placerat aliquam. Nunc pulvinar
massa et mattis lacinia.

\end{document}
\endinput
%%
%% End of file `sample-sigplan.tex'.

### 2023-07-02 Article.tex
Ny indledning
Ny indledning.

@@
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,anonymous,review]{acmart}
 \input{preamble}
-%%
-%% This is file `sample-sigplan.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigplan')
-%% 
-%% IMPORTANT NOTICE:
-%% 
-%% For the copyright see the source file.
-%% 
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigplan.tex.
-%% 
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%% 
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%%
-%% Commands for TeXCount
-%TC:macro \cite [option:text,text]
-%TC:macro \citep [option:text,text]
-%TC:macro \citet [option:text,text]
-%TC:envir table 0 1
-%TC:envir table* 0 1
-%TC:envir tabular [ignore] word
-%TC:envir displaymath 0 word
-%TC:envir math 0 word
-%TC:envir comment 0 0
-%%
-%%
-%% The first command in your LaTeX source must be the \documentclass
-%% command.
-%%
-%% For submission and review of your manuscript please change the
-%% command to \documentclass[manuscript, screen, review]{acmart}.
-%%
-%% When submitting camera ready or to TAPS, please change the command
-%% to \documentclass[sigconf]{acmart} or whichever template is required
-%% for your publication.
-%%
-%%
-
-%%
+
 %% \BibTeX command to typeset BibTeX logo in the docs
@@
 
-%% These commands are for a PROCEEDINGS abstract or paper.
 \acmConference[GPCE 2023]{Make sure to enter the correct
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{A generalised editor calculus (Short Paper)}
+\title{A type-safe generalized editor calculus (Short Paper)}
 
@@
 
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
 \begin{abstract}
@@
 \section{Introduction}
-This paper is on the topic of syntax-directed editors. 
-Most current examples of syntax-directed editor calculi, such as the ones presented by \cite{hazelnut} and \cite{type_safe_structure_editor}, are typically tailored specifically towards editing abstract syntax trees from an applied $\lambda$-calculus.
-This fact alone limits the usable scope of these calculi, since they would be rendered useless given any other form of abstract syntax.
-
-Furthermore, these calculi do not account for binding mechanisms in abstract syntax.
-Many of the editor calculi possess the ability to interleave construction and evaluation of programs, however, a consequence of this is a complex type system.
-
-In this paper we propose a generalized editor calculus that can be used to create a syntax-directed editor calculus for any abstract syntax. 
-Our editor calculus is inspired by higher-order abstract syntax presented in \cite{hoas}, and it will take binding-mechanisms into account.
-We have encoded our generalized editor calculus into a simply-typed lambda calculus extended with pairs, booleans, pattern matching and fixed-points.
-
-%Copy paste fra her
-\section{Generalized Editor Calculus}\label{sec:general_editor}
-We will now present a generalization of the syntax-directed editor calculus by \cite{type_safe_structure_editor}. 
-This specific editor calculus was designed with the language of an applied $\lambda$-calculus in mind. 
-%Instead, by using the notion of abstract syntax as presented in \cite{sec:pre_abstract_syntax} we generalize the concepts of the editor calculus to account for any language given by some abstract syntax.
-
-\subsection{Syntax}
-
-\subsubsection{Abstract Syntax}
-Firstly, we introduce the general representation of some abstract syntax that will be the focus of the editor calculus.
+
+Syntax-directed editors aka projectional editors are specialized
+editors that allow programmers to create programs directly using the
+syntax of the programming language. A prominent early example of an
+editor of this kind is that of the Cornell Program Synthesizer by Reps and
+Teitelbaum \cite{10.1145/358746.358755}, that was designed with the
+PL/I language in mind.
+
+More recent developments introduce models of computation that are
+specialized towards this form of editing. The Hazelnut framework by
+Omar et al. \cite{hazelnut} and the work by Godiksen et al. and
+\cite{type_safe_structure_editor} introduce editor calculi that
+describe the syntax and operational semantics of editing commands,
+including how such commands can be combined. This line of work has
+been tailored specifically towards editing abstract syntax trees from
+an applied $\lambda$-calculus.
+
+A major advantage of this approach is that it is typed -- each such editor calculus
+is equipped with a type system that guarantees that if an editor term
+$E$ is well-typed, then $e$ will produce a well-formed program.
+
+However, existing editor calculi do not account for context-sensitive
+aspects of program syntax, and notable not the binding 
+mechanisms that are prominenet in abstract syntax.
+
+Moreover, the work so far has been dependent on the syntax of the
+programming language that the editor must support. Whenever the
+programming language changes, the type safety results for the editor
+calculus have had to be proved anew. An example of this is the work by
+Andersen et al \cite{10.1145/3587216.3587221} on extending the editor
+calculus of \cite{type_safe_structure_editor} with let-declarations.
+
+In later work, Reps and Teitelbaum showed \cite{10.1145/390011.808247}
+how one can generalize their original work to build a Synthesizer
+Generator that would create a syntax-directed editor for any given
+programming language, given a specification of its syntax in the form
+of an attribute grammar. Around the same time, Donzeau-Gouge et
+al. developed the Mentor programming environment
+\cite{10.5555/800054.801990} that involved a command language, Mentol,
+and allowed designers to specify language-specific specializations
+using a specification language, Metal.
+
+In this paper we propose a generalized editor calculus that deals with
+these two issues.
+
+Our calculus deals with binding mechanisms by using notions from
+higher-order abstract syntax \cite{hoas}. Moreover, and importantly,
+we show how to encode encoded our generalized editor calculus into
+versions of the simply-typed lambda calculus extended with pairs,
+booleans, pattern matching and fixed-points. These fragments are all
+known to have type systems with strong safety guarantees, and in this
+way we a general result on type safety that applies to any instance of
+our generalized editor calculus.
+
+\section{A Generalized Editor Calculus}\label{sec:general_editor}
+
+We first present a generalization of the syntax-directed editor
+calculus by \cite{type_safe_structure_editor}.
+
+\subsection{Representing abstract syntax}
+
 We assume the abstract syntax is given by a set of sorts $\SORT$, an arity-indexed family of operators $\OP$ and a sort-indexed family of variables $\VAR$, such as presented in \cite{harper_foundations}.
-For the full definitions we refer to the full version of the paper.
-
-The notion of cursors and holes is central to the idea of an editor calculus.
-Whereas the calculus in \cite{type_safe_structure_editor} only has one cursor and hole term, in the general case we need a cursor and hole operator for every sort in the abstract syntax.
-With the set of sorts $\SORT$, the extended family of operators $\OP$ and family of variables $\VAR$ we can then define the sort-indexed family of abstract binding trees (refered to as "abt") $\ABT{}$ as the smallest family that satisfies the conditions of $\ABT{}$'s in \cite{harper_foundations}.
+
+The notion of cursors and holes is central to the idea of an editor
+calculus. Whereas the calculus in \cite{type_safe_structure_editor}
+only has one cursor and hole term, in the general case we need a
+cursor and hole operator for every sort in the abstract syntax.  With
+the set of sorts $\SORT$, the extended family of operators $\OP$ and
+family of variables $\VAR$ we can then define the sort-indexed family
+of abstract binding trees (refered to as "abt") $\ABT{}$ as the
+smallest family that satisfies the conditions of $\ABT{}$'s in
+\cite{harper_foundations}.
 
 \begin{example}\label{ex:abstract_syntax}
-    %To illustrate how the generalized editor calculus could be instantiated for some language, we will specialize an editor calculus as we proceed through this section for the abstract syntax shown in \cref{fig:example_abstract_syntax}.
-    To illustrate how the generalized editor calculus could be instantiated for some language, we will through our examples specialize an editor calculus.
-The abstract syntax for our specialization language is shown in \cref{fig:example_abstract_syntax}.
-    \federeboks{R C L L L}{
+Consider the following simple language of expressions with local declarations.
+\begin{align*}
   \text{Sort} && \text{Term} & \text{Operator} & \text{Arity} \\
@@
         & | & x & \text{var}[x] & ()e
-    }{\caption{Abstract syntax of arithmetic expressions and local declarations}\label{fig:example_abstract_syntax}}
-    The syntax in \cref{fig:example_abstract_syntax} describes a very simple language of statements and expressions. For example, using the abstract syntax we could construct the statement
+\end{align*}
+ We can here represent
     \begin{equation*}
@@
     \end{equation*}
-    which, using operators, would be represented as the abt
+    as the abt
     \begin{equation*}
@@
     \end{equation*} 
-
-    Given the abstract syntax in \cref{fig:example_abstract_syntax} we can extend it with holes and cursors. This gives us the new operators shown in \cref{fig:example_holes_and_cursors}.
-    \federeboks{R C L L L}{
+%
+    We can extend this language with holes and cursors and get
+    %
+\begin{align*}
         \text{Sort} && \text{Term} & \text{Operator} & \text{Arity} \\
@@
         & | & \hole{e} & \text{hole}_e & ()e
-    }{\caption{Introduction of cursors and holes to the abstract syntax in \cref{fig:example_abstract_syntax}}\label{fig:example_holes_and_cursors}}
-    With this extension we can now represent statements with holes and cursors as abt's. For example the statement
+ \end{align*}
+%
+    With this extension the statement
     \begin{equation*}
@@
     \end{equation*}
-    would be represented as the abt
+    is represented as the abt
     \begin{equation*}
@@
 
-\subsubsection{Editor Calculus}
-The abstract syntax of the generalized editor calculus is shown in \cref{fig:editor_calculus_syntax}. It closely resembles that of the calculus presented by \cite{type_safe_structure_editor}.
-\federeboks{R C L}{
-    E & ::= & \pi.E \ | \ \phi => E_1|E_2 \ | \ E_1 \ggg E_2 \ | \ \text{rec} \ x.E \ | \ x \ | \ \text{nil} \\
-    \pi & ::= & \text{child} \ n \ | \ \text{parent} \ | \ \{o\} \\
-    \phi & ::= & \neg\phi \ | \ \phi_1 \wedge \phi_2 \ | \ \phi_1 \vee \phi_2 \ | \ @o \ | \ \Diamond o \ | \ \square o
-}{\caption{Abstract syntax of general editor calculus}\label{fig:editor_calculus_syntax}}
-Editor expressions $E \in Edt$, where $Edt$ is the sort of editor expressions, describes the behaviour of the editor calculus.
-The semantics behind the different constructs are defined in section \cref{sec:Semantics}.
-
+\subsubsection{The editor calculus}
+Editor expressions $E \in \Edt$, where $\Edt$ is the sort of editor
+expressions, describe the possible edit actions. The formation rules
+are given below.
+%
+\begin{align*}
+  E & ::= & \pi.E  \mid  \phi => E_1|E_2  \mid  E_1 \ggg E_2  \mid  \text{rec} \ x.E  \mid  x  \mid  \text{nil} \\
+    \pi & ::= & \text{child} \ n  \mid  \text{parent}  \mid  \{o\} \\
+    \phi & ::= & \neg\phi  \mid  \phi_1 \wedge \phi_2  \mid  \phi_1 \vee \phi_2  \mid  @o  \mid  \Diamond o  \mid  \square o
+\end{align*}
+%
 Evaluation of abt's, which \cite{type_safe_structure_editor} introduces through a $eval$ construct, is not included in our generalized version.
@@
 \begin{example}\label{ex:editor_calculus}
-    Expanding upon \cref{ex:abstract_syntax}, we can specialize the abstract syntax of the general editor calculus presented in \cref{fig:editor_calculus_syntax} for our simple language by defining the set of operators $o$ ranges over.
-In this case we have that
+    We can define a version of the editor calculus for
+    \cref{ex:abstract_syntax} by defining the set of operators $o$ ranges over. In this case we have that
     \begin{equation*}
@@
     \end{equation*}
-    allowing us to write editor expressions such as $(@\text{hole}_e => \{\text{plus}\}.\text{nil}|\text{nil})$, which would substitute the current tree encapsulated by the cursor with the $plus$ operator, if the cursor is at the operator $\text{hole}_e$.
+    We can then write e.g. $(@\text{hole}_e =>
+    \{\text{plus}\}.\text{nil}|\text{nil})$. This expression would
+    substitute the current tree encapsulated by the cursor with the
+    $plus$ operator, if the cursor is at the operator
+    $\text{hole}_e$. 
 \end{example}
 
-\subsubsection{Cursor Context}
+\subsubsection{Cursor contexts}
 In editor expressions, we frequently refer to performing actions on, or with, the abt that is currently encapsulated by the cursor.
@@
 \subsubsection{Editor Expressions} \label{sec:Semantics}
-The labelled transition system for editor expressions is defined as $(Edt \times \ABT{}, Apc \cup \{\epsilon\}, \Rightarrow)$.
-Transitions are of the form $\lr{E, a} \ltrans{\alpha} \lr{E', a'}$ where the editor expression $E \in Edt$ is closed and the abt $a \in \ABT{}$ is well-formed.
-The labels of the transitions, $\alpha$, are the atomic prefix commands $Apc$ and the silent transition $\epsilon$.
-The transition relation $\Rightarrow$ is defined by the rules shown in \cref{fig:editor_expression_rules}.
-These facilitate ways of composing editor expressions together with conditions, sequential composition and recursion.
-The means of traversing and modifying an abt are provided by the prefixed expression $\pi.E$, which evaluates $\pi$ before continuing with $E$.
-The conditional expression $\phi => E_1|E_2$ reduces to $E_1$ if $\phi$ is satisfied and $E_2$ otherwise.
-The sequential expression $E_1 \ggg E_2$ evaluates $E_2$ only once $E_1$ has been reduced to $nil$.
-The recursive expression $rec \ x.E$ binds the recursion variable $x$ in $E$, which can then be used to recursively iterate an expression.
+
+The labelled transition system for editor expressions is defined as
+$(\Edt \times \ABT{}, Apc \cup \{\epsilon\},
+\Rightarrow)$. Transitions are of the form $\lr{E, a} \ltrans{\alpha}
+\lr{E', a'}$ where the editor expression $E \in \Edt$ is closed and
+the abt $a \in \ABT{}$ is well-formed. The labels of the transitions,
+$\alpha$, are the atomic prefix commands $Apc$ and the silent
+transition $\epsilon$. The transition relation $\Rightarrow$ is
+defined by the rules shown in
+\cref{fig:editor_expression_rules}. These facilitate ways of composing
+editor expressions together with conditions, sequential composition
+and recursion. The means of traversing and modifying an abt are
+provided by the prefixed expression $\pi.E$, which evaluates $\pi$
+before continuing with $E$. The conditional expression $\phi =>
+E_1|E_2$ reduces to $E_1$ if $\phi$ is satisfied and $E_2$
+otherwise. The sequential expression $E_1 \ggg E_2$ evaluates $E_2$
+only once $E_1$ has been reduced to $nil$. The recursive expression
+$rec \ x.E$ binds the recursion variable $x$ in $E$, which can then be
+used to recursively iterate an expression.
 
@@
 
-%A concrete overview of the calculus and the constructs necessary to encode every central concept of the editor calculus can be seen in \cref{tab:encoding_requirements}.
-
-%\begin{table}[H]
-%    \centering
-%    \begin{tabular}{| l | l |}
-%        \hline
-%        Editor calculus & Encoding\\
-%        \hline
-%        \ Abstract Binding Trees & \ Simply typed lambda calculus, \stlc \\
-%        \hline
-%        \ Atomic Prefix Commands & \ \stlc with pattern matching, $\lambda_{->,p}$\\
-%        \hline
-%        \ Modal Logic & \ \stlc with pattern matching and recursion, $\lambda_{->,p,fix}$ \\
-%        \hline
-%        \ Editor Expressions & \ \stlc with pattern matching and recursion, $\lambda_{->,p,fix}$ \\
-%        \hline 
-%    \end{tabular}
-%    \vspace{1mm}
-%    \caption{Calculus requirements to encode the generalized editor calculus}
-%    \label{tab:encoding_requirements}
-%\end{table}
+A concrete overview of the calculus and the constructs necessary to encode every central concept of the editor calculus can be seen in \cref{tab:encoding_requirements}.
+
+\begin{table*}
+   \centering
+   \begin{tabular}{| l | l |}
+       \hline
+       \textsf{Editor calculus} & \textsf{Encoding}\\
+       \hline\hline
+       \ Abstract Binding Trees & \ Simply typed lambda calculus, \stlc \\
+       \hline
+       \ Atomic Prefix Commands & \ \stlc with pattern matching, $\lambda_{->,p}$\\
+       \hline
+       \ Modal Logic & \ \stlc with pattern matching and recursion, $\lambda_{->,p,fix}$ \\
+       \hline
+       \ Editor Expressions & \ \stlc with pattern matching and recursion, $\lambda_{->,p,fix}$ \\
+       \hline 
+   \end{tabular}
+   \vspace{1mm}
+   \caption{Editor calculus fragments and the lambda calculus fragments
+     used for encoding them}
+   \label{tab:encoding_requirements}
+\end{table*}
 
@@
 We extend the lambda-calculus with term constants $o$ for every $o \in \OP$ excluding cursors, and the base types $s$ for every $s \in \SORT$.
-Typing rules for operators can be seen on \cref{fig:abt_typing_rules}, where we can infer the type of an operator in the lambda-calculus by its arity. 
+Typing rules for operators can be seen on \cref{fig:abt_typing_rules},
+where we can infer the type of an operator in the lambda-calculus by
+its arity.  
 \federeboks{C}{
@@
 
-With the operators added as term constants and sorts added as base types, encoding abts is straightforward as we just have to curry the operator $o$, encode the children $a_1, \dots, a_n$ and add typings.
-This can be seen on \cref{fig:enc_abt}.
+With the operators added as term constants and sorts added as base
+types, encoding abts is straightforward as we just have to curry the
+operator $o$, encode the children $a_1, \dots, a_n$ and add
+typings. This can be seen on \cref{fig:enc_abt}.
 
@@
 %% End of file `sample-sigplan.tex'.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2023-07-02 Article.tex
Mere nyt

@@
 Therefore, it becomes necessary to introduce the concept of well-formed abt's.
-We will do this informally, where a well-formed abt is simply an abt containing exactly one cursor.
+Informally, a well-formed abt is simply an abt containing exactly one cursor.
 This is guarantied when an abt can be interpreted as $C[\dot{a}]$, meaning the context C, where $\ctxhole$ is substituted with a tree $\dot{a}$, where the cursor-operator is either the root $\dot{a}$, or one of the children of the root.
@@
 
-\begin{example}\label{ex:well_formed_trees}
-   We will now show the extension of \cref{fig:example_abstract_syntax}, where cursors are included, as shown in \cref{fig:example_well_formed_trees}.
+% \begin{example}\label{ex:well_formed_trees}
+%    We will now show the extension of \cref{fig:example_abstract_syntax}, where cursors are included, as shown in \cref{fig:example_well_formed_trees}.
 We note that $\hat{s}$ and $\hat{e}$ are trees without cursors.
-    \federeboks{R C l}{
-        \dot{s} & ::= & let $x = [\hat{e}]$ in $\hat{s}$ | let $x = \hat{e}$ in $[\hat{s}]$ | $[\hat{e}]$ | $[\hat{s}]$ \\
-        \dot{e} & ::= & $[\hat{e}_1] + \hat{e}_2$ | $\hat{e}_1 + [\hat{e}_2]$ | $[\hat{e}]$
-    }{\caption{Abstract syntax of abt's with exactly one cursor}\label{fig:example_well_formed_trees}}
-    Given the syntax in \cref{fig:example_well_formed_trees} and the previously introduced cursor context, we can now determine the well-formedness of any abt in our language.
+%     \federeboks{R C l}{
+%         \dot{s} & ::= & let $x = [\hat{e}]$ in $\hat{s}$ | let $x = \hat{e}$ in $[\hat{s}]$ | $[\hat{e}]$ | $[\hat{s}]$ \\
+%         \dot{e} & ::= & $[\hat{e}_1] + \hat{e}_2$ | $\hat{e}_1 + [\hat{e}_2]$ | $[\hat{e}]$
+%     }{\caption{Abstract syntax of abt's with exactly one cursor}\label{fig:example_well_formed_trees}}
+%     Given the syntax in \cref{fig:example_well_formed_trees} and the previously introduced cursor context, we can now determine the well-formedness of any abt in our language.
 For example the statement
-    \begin{equation*}
-        \text{let $x=[\hole{e}]$ in $x+5$}
-    \end{equation*}
-    is well-formed since it can be written as $C[\dot{e}]$ where
-    \begin{align*}
-        C &= \text{let $x=\ctxhole$ in $x+5$} \\
-        \dot{e}&= [\hole{e}]
-    \end{align*}
-    It should be noted that there often are two valid interpretations of a tree $a$ as $C[\dot{a}]$. In this example, the other valid interpretation would be:
-    \begin{align*}
-        C' &= \ctxhole \\
-        \dot{e}' &= \text{let $x=[\hole{e}]$ in $x+5$}
-    \end{align*}
-\end{example}
+%     \begin{equation*}
+%         \text{let $x=[\hole{e}]$ in $x+5$}
+%     \end{equation*}
+%     is well-formed since it can be written as $C[\dot{e}]$ where
+%     \begin{align*}
+%         C &= \text{let $x=\ctxhole$ in $x+5$} \\
+%         \dot{e}&= [\hole{e}]
+%     \end{align*}
+%     It should be noted that there often are two valid interpretations of a tree $a$ as $C[\dot{a}]$. In this example, the other valid interpretation would be:
+%     \begin{align*}
+%         C' &= \ctxhole \\
+%         \dot{e}' &= \text{let $x=[\hole{e}]$ in $x+5$}
+%     \end{align*}
+% \end{example}
 
@@
 
-Unlike the editor calculus by \cite{type_safe_structure_editor}, we do not utilize any form of structural congruence on editor expressions to define the semantics of trivial sequential composition or recursion.
-Instead we simply define these notions with their own transition rule (Seq-Trivial) and (Recursion)
+% Unlike the editor calculus by \cite{type_safe_structure_editor}, we do
+% not utilize any form of structural congruence on editor expressions to
+% define the semantics of trivial sequential composition or
+% recursion. Instead we simply define these notions with their own
+% transition rule (Seq-Trivial) and (Recursion) 
 \federeboks{c}{
@@
 
-\section{Encoding the generalized editor calculus in an extended lambda calculus} \label{sec:encoding}
-In this section we attempt to soundly encode the generalized editor calculus using the simply typed $\lambda$-calculus (\stlc) as the base calculus.
-The calculus will be extended with various constructs such as pattern matching, pairs and the recursive \textit{fix} operator.
-To encode the abstract binding trees, we will only need the simply typed lambda calculus.
-To encode the contexts, we introduce pairs.
-To encode atomic prefix commands we further extend the calculus with pattern matching ($\lambda_{->,p}$).
-Lastly to encode modal logic and editor expressions we extend the calculus with a fixed point operator ($\lambda_{->,p,fix}$).
-
-A concrete overview of the calculus and the constructs necessary to encode every central concept of the editor calculus can be seen in \cref{tab:encoding_requirements}.
+\section{Encoding the generalized editor calculus} \label{sec:encoding}
+
+To encode the abstract binding trees, the simply typed lambda calculus
+suffices.
+
+In order to encode the generalized editor calculus using the simply
+typed $\lambda$-calculus (\stlc) as the base calculus, we extend the
+latter with new constructs.
+
+To encode the contexts, we introduce pairs. To encode atomic prefix
+commands we further extend the calculus with pattern matching
+($\lambda_{->,p}$). Finally, to encode modal logic and editor
+expressions recursion is needed in the form of the fixed point
+operator ($\lambda_{->,p,fix}$).
+
+An overview of the calculus and the constructs necessary to encode
+each part of it can be seen in \cref{tab:encoding_requirements}.
 
@@
 
-\subsection{Motivation}
-The motivation behind encoding the generalized editor calculus in $\lambda_{->,p,fix}$ (or subsets thereof) is most importantly the type system it provides.
-If our encoding is sound, then any instance of the editor calculus will have a sound type system, regardless of the abstract syntax.
-This follows from the fact that the type system of $\lambda_{->,p,fix}$ is sound.
-Our editor calculus having a sound type system means that it will reject incorrect editor expressions.
-
-Secondly a $\lambda_{->,p,fix}$ encoding provides a clear strategy for implementing any instantiation of the generalized editor calculus in a functional programming language.
-
-\subsection{Abstract Binding Trees}
+ All of the lambda calculi that we consider in the following have a
+ standard reduction semantics with reductions of the form $e \to
+ e'$. Moreover, every calculus has a type systems with type judgements
+ of the form $\Gamma \vdash e : \tau$, where $\tau$ ranges over a set
+ of types \Types\ and $\Gamma$ is a type environment
+ $\Gamma : \Var \pra \Types$. The type systems all satisfy the subject
+ reduction property.
+
+ \begin{theorem}[Subject reduction]\label{thm:subred}
+   If $\Gamma \vdash e : \tau$ and $e \to e'$, then $\Gamma \vdash e'
+   : \tau$.
+ \end{theorem}
+
+ It is this result that ensures that our encoding will indeed provide
+ a general result about type soundness for the generalized editor
+ calculus.
+
+\subsection{Abstract binding trees}
 To encode abstract binding trees we only need to use the simply typed lambda-calculus.
@@
     }{
-        \caption{Types for the encoded operators.}
+        \caption{Types of the encoded operators.}
         \label{fig:abt_example_type}
@@
 
-\subsection{Cursor Contexts}\label{sec:enc_cursor_contexts}
+\subsection{Cursor contexts}\label{sec:enc_cursor_contexts}
 
@@
 
-\subsection{Atomic Prefix Commands}
-To encode the atomic prefix commands we further extend the lambda calculus with pattern matching.
-As seen in \cref{fig:apc_syntactic_forms} we introduce the match construct where we match the term $M$ on the patterns $p$ resulting in the corresponding term $N$ if there is a match.
-As mentioned $p$ is the syntactic category describing the pattern we are trying to match.
-These patterns consist of variables, wildcards, operators, pairs, and lastly bindings which recursively match on the body of an abstraction.
-\federeboks{R C L l}{
-    \multicolumn{4}{c}{Terms}
+\subsection{Atomic prefix commands}
+
+To encode the atomic prefix commands we further extend the lambda
+calculus with pattern matching. The formation rules are extended as
+shown below; $p$ is the metavariable for patterns.
+% As seen in
+% \cref{fig:apc_syntactic_forms} we introduce the match construct where
+% we match the term $M$ on the patterns $p$ resulting in the
+% corresponding term $N$ if there is a match. As mentioned $p$ is the
+% syntactic category describing the pattern we are trying to
+% match. These patterns consist of variables, wildcards, operators,
+% pairs, and lastly bindings which recursively match on the body of an
+% abstraction.  
+% \federeboks{R C L l}{
+%     \multicolumn{4}{c}{Terms}
+%     \\
+\begin{align*}
+    M, N & ::= & \text{match}\ M\ \overrightarrow{p \to N} && \text{(match construct)}
     \\
-    M, N & ::= & \text{match}\ M\ \overrightarrow{p \to N} & (match construct)
-    \\
-    p & ::= & x              & (variable) \\
-      & |   & \_             & (wildcard) \\
-      & |   & o\ \Vec{p}     & (operator) \\
-      & |   & (p_1, p_2)     & (pair) \\
-      & |   & .p             & (binding)
-}{\caption{Abstract syntax for representing atomic prefix commands, extending the lambda-calculus with a match construct.}\label{fig:apc_syntactic_forms}}
-The reduction rules for the match construct can be seen in the full version of the paper.
-
-We now define the auxiliary functions as seen in \cref{tab:zip_oper}.
-These function as the operations of a zipper data structure, which helps encode atomic prefix commands on abts.
-We also introduce the abbreviation $[a]$ to mean the cursor operator with the child $a$.
+    p & ::= & x              & & \text{(variable)} \\
+      & |   & \_             && \text{(wildcard)} \\
+      & |   & o\ \Vec{p}     && \text{(operator)} \\
+      & |   & (p_1, p_2)     && \text{(pair)} \\
+         & |   & .p             && \text{(binding)}
+\end{align*}
+%} {\caption{Abstract syntax for representing atomic prefix commands, extending the lambda-calculus with a match construct.}\label{fig:apc_syntactic_forms}}
+%The reduction rules for the match construct can be seen in the full version of the paper.
+%
+We are now able to define functions functions that describe navigation
+in an abt and substitution; the definitions are givren in
+\cref{tab:zip_oper}. We introduce the abbreviation $[a]$ to mean the
+cursor operator with the child $a$.
 
 \federeboks{L}{
-    $down$ \define \lambda x:s.\text{match }x
+    \textit{down} \define \lambda x:s.\text{match }x
     \quad [o \ (.a_1) \dots (.a_n)] \to \\
@@
 
-    $right$ \define \lambda x:s.\text{match }x
+    \textit{right} \define \lambda x:s.\text{match }x
     \quad o \ (.a_1) \ \dots (.[a_i]) \ \dots (.a_n) \to \\
@@
 
-    $up$ \define \lambda x:s.\text{match }x
+    \textit{up} \define \lambda x:s.\text{match }x
     \quad o \ (.a_1) \dots (.[a_i]) \dots (.a_n) \to \\
@@
 
-    $set$ \define \lambda a:s.\lambda x:s.\text{match }x
+    \textit{set} \define \lambda a:s.\lambda x:s.\text{match }x
     \quad [a'] \to [a]
-}{\caption{Function definitions for cursor movement and substitution on abstract binding trees.}\label{tab:zip_oper}}
-The encoding of atomic prefix commands can be seen in \cref{fig:atc_enc}.
-The $child$ command can be encoded as a base case and a recursive case.
-In the base case we want to move to the first $child$, which means we just go down.
-In the recursive case we want to move to $child$ $n$, we can encode this as moving to $child$ $n-1$ and then moving right.
-$parent$ and $insert$ can directly be encoded as $up$ and $set$ respectively.
+}{\caption{Defining cursor movement and
+    substitution.}\label{tab:zip_oper}}
+
+The encoding of atomic prefix commands are given in
+\cref{fig:atc_enc}. Note that the $child$ command is defined
+recursively. \textit{parent} and \textit{insert} are simply encoded as \textit{up} and \textit{set} respectively.
 \federeboks{L L}{
-    \encode{\text{child} \ 1} = \text{down}
+    \encode{\textit{child} \ 1} = \textit{down}
     &
-    \encode{\text{child} \ n} = \text{right}\ \encode{\text{child} \ n - 1}
-    \\\\
-    \encode{\text{parent}} = \text{up}
+    \encode{\textit{child} \ n} = \textit{right}\ \encode{\text{child} \ n - 1}
+    \\\\
+    \encode{\textit{parent}} = \textit{up}
     &
-    \encode{\text{insert} \ a} = \text{set} \ \encode{a}
+    \encode{\textit{insert} \ a} = \textit{set} \ \encode{a}
 }{
@@
 
-\subsection{Control Structures and Modal Logic}
-To encode control structures and modal logic, we need to further extend the lambda-calculus with a \fix operator as well as the $Bool$ base type.
-We introduce the \fix operator as seen in \cite{types_programming_languages}, as well as boolean types and values.
-On \cref{fig:edt_enc} the encoding of editor expressions can be seen.
-For atomic prefixes the atomic prefix command should be applied before the rest of the editor expression.
-The $nil$ editor expression will become the identity function.
-Sequential editor expressions will apply the encoding of the editor expressions in the correct order.
-Recursion works by using the newly introduced \fix operator.
-Lastly, conditional editor expressions are encoded such that we match on the evaluation of $\phi$ on the first projection of a context $C$ and apply the encoding of $E_1$ or $E_2$ on $C$ accordingly.
-For the full encoding of modal logic, we refer to the full version of the paper.
+\subsection{Control structures and modal logic}
+
+To encode control structures and modal logic, we need to further
+extend the lambda-calculus with a \fix operator as well as the $Bool$
+base type. We introduce the \fix operator as seen in
+\cite{types_programming_languages}, as well as boolean types and
+values.
+
+\cref{fig:edt_enc} shows the encoding of editor expressions. For
+atomic prefixes the atomic prefix command is applied before the
+rest of the editor expression. The $nil$ editor expression is
+the identity function. Sequential editor expressions apply the
+encoding of the editor expressions in the correct order. Recursion
+works by using the \fix operator. Lastly, conditional
+editor expressions are encoded such that we match on the evaluation of
+$\phi$ on the first projection of a context $C$ and apply the encoding
+of $E_1$ or $E_2$ on $C$ accordingly. 
 %\federeboks{R C L l}{
@@
 }
-The encoding of conjunction, disjunction and negation, shown in \cref{fig:encoding_connectives}, comes quite naturally using the previously defined auxiliary functions.
+The encoding of propositional connectives is shown in
+\cref{fig:encoding_connectives}.
+
 \federeboks{L}{
@@
 \fi
-With this, the entire editor calculus has been encoded. We refer to the full version of the paper to see the proof of the soundness of our encoding.
-\iffalse
-\begin{example}
-    To illustrate how editor expressions are encoded, we provide an encoding of the following expression
-    \begin{equation*}
-        \text{child\ $2$.nil $\ggg$ parent.nil}
-    \end{equation*}
-    using the encoding rules seen on \cref{fig:edt_enc} and \cref{fig:atc_enc}. For readability purposes we will not add explicit typings, just note that any variable $C_i$ has type $Ctx$.
-    \begin{align*}
-         &\ \encode{\text{child\ $2$.nil $\ggg$ parent.nil}} \\
-         =&\ \lambda C_1 . \encode{\text{parent.nil}} (\encode{\text{child\ $2$.nil}} C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . \encode{\text{nil}}(\encode{\text{parent}} C_2.1, C_2.2)) (\encode{\text{child\ $2$.nil}} C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . \encode{nil}(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . (\lambda C_3 . C_3)(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
-         \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . \encode{nil} (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\lambda C_4.C_4) (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
-         \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (\encode{child\ 1} (C_3.1, C_3.2))C_1) \\
-         =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (down (C_3.1, C_3.2))C_1) \\
-    \end{align*}
-    As can be seen, the editor expression could be fully encoded to the extended lambda calculus. We refer to the full paper for
-\end{example}
-\fi
-
-\section{Conclusion \& Further Work}
-%Conclusion
-We have developed a generalized editor calculus that enables the creation of a syntax-directed editor calculus for a specific abstract syntax.
-Subsequently, we encoded this editor calculus into an extended version of the lambda calculus, which incorporates pairs, recursion and pattern matching.
-Further work on our generalized editor calculus could be to prove whether the encoding is complete.
-Otherwise, one could implement a redo-undo construct such as introduced in \cite{undo_Edit} in the generalized editor calculus.
-Another interesting construct to implement would be a copy-paste operation.
+
+% \begin{example}
+%     We are now able to give an example of how to encode editor
+%     expressions. Consider
+%     \begin{equation*}
+%         \text{child\ $2$.nil $\ggg$ parent.nil}
+%     \end{equation*}
+%     using the encoding rules seen on \cref{fig:edt_enc} and \cref{fig:atc_enc}. For readability purposes we will not add explicit typings, just note that any variable $C_i$ has type $Ctx$.
+%     \begin{align*}
+%          &\ \encode{\text{child\ $2$.nil $\ggg$ parent.nil}} \\
+%          =&\ \lambda C_1 . \encode{\text{parent.nil}} (\encode{\text{child\ $2$.nil}} C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . \encode{\text{nil}}(\encode{\text{parent}} C_2.1, C_2.2)) (\encode{\text{child\ $2$.nil}} C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . \encode{nil}(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . (\lambda C_3 . C_3)(up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
+%          \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\encode{child\ 2.nil} C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . \encode{nil} (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\lambda C_4.C_4) (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
+%          \betar&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . (\encode{child\ 2} C_3.1, C_3.2)C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (\encode{child\ 1} (C_3.1, C_3.2))C_1) \\
+%          =&\ \lambda C_1 . (\lambda C_2 . (up\ C_2.1, C_2.2)) (\lambda C_3 . right (down (C_3.1, C_3.2))C_1) \\
+%     \end{align*}
+%     As can be seen, the editor expression could be fully encoded to the extended lambda calculus. We refer to the full paper for
+% \end{example}
+
+\subsection{Soundness of the encoding}
+
+The main result about our encoding is that it is sound wrt. the
+operational semantics of the original calculus.
+
+\begin{theorem}[Soundness] For every editor expression $E$ and abt $a$
+  we have that if
+  %
+  \[ \lr{E,a} \ltrans{} \lr{E',a'} \]
+  then
+  \[ \encode{E}\encode{a} \rightarrow^{\ast}  \encode{E'}\encode{a'}  \]
+\end{theorem}
+
+\section{Conclusion and Further Work}
+
+We have developed a generalized editor calculus that enables the
+creation of a syntax-directed editor calculus for a specific abstract
+syntax. Subsequently, we encoded this editor calculus into an extended
+version of the lambda calculus, which incorporates pairs, recursion
+and pattern matching.  Further work on our generalized editor calculus
+could be to prove whether the encoding is complete. Otherwise, one
+could implement a redo-undo construct such as introduced in
+\cite{undo_Edit} in the generalized editor calculus. Another
+interesting construct to implement would be a copy-paste operation.
 
@@
 
-%We have identified several possibilities for expanding our work on the generalized editor calculus.
-First of all, it would be excellent to prove that our encoding is complete.
-However, in the scope of this paper, we decided to only include the proof of the soundness of the encoding, as that would give us the sound type system.
+We have identified several possibilities for expanding our work on the
+generalized editor calculus. First of all, the encoding should be
+proved complete. Secondly,
 

### 2023-07-04 Article.tex
Ny udgave

@@
   2023}{Woodstock, NY}
-%%
-%%  Uncomment \acmBooktitle if the title of the proceedings is different
-%%  from ``Proceedings of ...''!
-%%
-%%\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-%%  June 03--05, 2018, Woodstock, NY}
+
 \acmPrice{15.00}
@@
    <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+       <concept_id>10011007.10011006.10011050.10011023</concept_id>
+       <concept_desc>Software and its engineering~Specialized application languages</concept_desc>
        <concept_significance>500</concept_significance>
@@
    <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
+       <concept_id>10003752.10010124.10010131</concept_id>
+       <concept_desc>Theory of computation~Program semantics</concept_desc>
+       <concept_significance>500</concept_significance>
        </concept>
    <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+       <concept_id>10003752.10010124.10010131.10010134</concept_id>
+       <concept_desc>Theory of computation~Operational semantics</concept_desc>
+       <concept_significance>500</concept_significance>
        </concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Software and its engineering~Specialized application languages}
+\ccsdesc[500]{Theory of computation~Program semantics}
+\ccsdesc[500]{Theory of computation~Operational semantics}
 
@@
 
-%To ensure that the cursor can only reside somewhere within the subtree specified by the context hole, we need to introduce some notion of cursorless abt's, written as $\hat{s}$.
-Using the abstract syntax, given by $\SORT$, $\OP$ and $\VAR$, we define a new corresponding abstract syntax which can only build abt's without cursors.
-
-%The abstract syntax of cursor contexts are defined according to \cref{def:cursor_contexts}.
-Using cursor contexts we can interpret any sort of abt $a \in \ABT{}$ as the cursor context $C[a']$, where the abt $a'$ is substituted for the context hole $\ctxhole$.
-Due to \cref{def:cursor_contexts} this also means that the cursor can only reside within $a'$, since it cannot be anywhere in $C$.
-
-%\begin{definition}[Cursor Contexts]\normalfont
-%    \label{def:cursor_contexts}
- %   The abstract syntax of cursor contexts is defined as an extension of the syntax for cursorless abt's. It is given by
-%    \begin{enumerate}
-%        \item The sorts $\SORT^C = \SORTc \cup \{ C \}$
-%        \item The family of operators $\OP^C = \OPc$ extended with the $\ctxhole$ operator of arity $()C$.
- %       \item Furthermore, for every operator $\hat{o} \in \OPc$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_n}.\hat{s}_n)\hat{s}$ and for every $1 \leq i \leq n$ we also add the operator $o^C_i$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_i}.C, \dots, \Vec{\hat{s}_n}.\hat{s}_n)C$ to $\OP^C$.
-These represents that $\ctxhole$ must be somewhere in subtree $i$ of operator $\hat{o}$.
-%        \item The family of variables $\VAR^C = \VARc$
-%    \end{enumerate}
-%\end{definition}
-
-%\begin{example}\label{ex:cursor_contexts}
-%    Continuing from \cref{ex:editor_calculus}, we will now extend the editor calculus for our simple language with cursorless abt's and cursor contexts.
-
- %   Firstly, we define the abstract syntax of abt's without cursors, the result of which is shown in \cref{fig:example_cursorless_trees}.
-For brevity the syntax is presented as formation rules, since translating these into the notion of operators, while trivial, is not a very concise way of presenting the abstract syntax.
- %   \federeboks{R C l}{
-  %      \hat{s} & ::= & let $x = \hat{e}$ in $\hat{s}$ | %$\hat{e}$ | $\hole{s}$ \\
-  %      \hat{e} & ::= & $\hat{e}_1 + \hat{e}_2$ | n | x | $\hole{e}$
- %   }{\caption{Abstract syntax of abt's without cursors}\label{fig:example_cursorless_trees}}
- %   Secondly, using the syntax of cursorless abt's we can define the abstract syntax of cursor contexts according to \cref{def:cursor_contexts}. This is shown in \cref{fig:example_cursor_contexts}.
- %   \federeboks{R C l}{
- %       C & ::= & let $x = C$ in $\hat{s}$ | let $x = \hat{e}$ in $C$ \\
-  %      & | & $C + \hat{e}_2$ | $\hat{e}_1 + C$ | $\ctxhole$
- %   }{\caption{Abstract syntax of cursor contexts}\label{fig:example_cursor_contexts}}
-%\end{example}
-
 Although a cursor context accurately specifies the subtree where the cursor must reside, it does not inherently guarantee the presence of only one cursor within that subtree.
@@
 This is guarantied when an abt can be interpreted as $C[\dot{a}]$, meaning the context C, where $\ctxhole$ is substituted with a tree $\dot{a}$, where the cursor-operator is either the root $\dot{a}$, or one of the children of the root.
-
-%\begin{definition}[Well-formed trees]\normalfont
-%    \label{def:well_formed_trees}
-%    Firstly, we define an abstract syntax that describes abt's with exactly one cursor either as the root or as an argument of the root. The syntax is given by
- %   \begin{enumerate}
- %       \item The sorts $\SORTcc = \SORTc \cup \{ \dot{s} \}_{s \in \SORT}$.
- %       \item The family of operators $\OPcc = \OPc$ extended with an operator of arity $(\hat{s})\dot{s}$ for every $\hat{s} \in \SORTc$.
-These represent that the cursor encapsulates the root of a cursorless abt of sort $\hat{s}$.
-   %     \item Furthermore, for every operator $\hat{o} \in \OPc$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_n}.\hat{s}_n)\hat{s}$ and for every $1 \leq i \leq n$ we also add the operator $\dot{o}_i$ of arity $(\Vec{\hat{s}_1}.\hat{s}_1, \dots, \Vec{\hat{s}_i}.\hat{s}_i, \dots, \Vec{\hat{s}_n}.\hat{s}_n)\dot{s}$ to $\OPcc$.
-These represent that the cursor encapsulates the cursorless subtree $i$ of operator $\hat{o}$.
-%        \item The family of variables $\VARcc = \VARc$.
- %   \end{enumerate}
-%    Given this, a well-formed abt $a \in \ABT{}$ is any abt that can be interpreted as $C[\dot{a}]$, since $C$ cannot contain any cursors and $\dot{a}$ contains exactly one.
-%\end{definition}
-
-% \begin{example}\label{ex:well_formed_trees}
-%    We will now show the extension of \cref{fig:example_abstract_syntax}, where cursors are included, as shown in \cref{fig:example_well_formed_trees}.
-We note that $\hat{s}$ and $\hat{e}$ are trees without cursors.
-%     \federeboks{R C l}{
-%         \dot{s} & ::= & let $x = [\hat{e}]$ in $\hat{s}$ | let $x = \hat{e}$ in $[\hat{s}]$ | $[\hat{e}]$ | $[\hat{s}]$ \\
-%         \dot{e} & ::= & $[\hat{e}_1] + \hat{e}_2$ | $\hat{e}_1 + [\hat{e}_2]$ | $[\hat{e}]$
-%     }{\caption{Abstract syntax of abt's with exactly one cursor}\label{fig:example_well_formed_trees}}
-%     Given the syntax in \cref{fig:example_well_formed_trees} and the previously introduced cursor context, we can now determine the well-formedness of any abt in our language.
-For example the statement
-%     \begin{equation*}
-%         \text{let $x=[\hole{e}]$ in $x+5$}
-%     \end{equation*}
-%     is well-formed since it can be written as $C[\dot{e}]$ where
-%     \begin{align*}
-%         C &= \text{let $x=\ctxhole$ in $x+5$} \\
-%         \dot{e}&= [\hole{e}]
-%     \end{align*}
-%     It should be noted that there often are two valid interpretations of a tree $a$ as $C[\dot{a}]$. In this example, the other valid interpretation would be:
-%     \begin{align*}
-%         C' &= \ctxhole \\
-%         \dot{e}' &= \text{let $x=[\hole{e}]$ in $x+5$}
-%     \end{align*}
-% \end{example}
 
@@
 
-% Unlike the editor calculus by \cite{type_safe_structure_editor}, we do
-% not utilize any form of structural congruence on editor expressions to
-% define the semantics of trivial sequential composition or
-% recursion. Instead we simply define these notions with their own
-% transition rule (Seq-Trivial) and (Recursion) 
 \federeboks{c}{
@@
 \subsubsection{Substitution and cursor movement}
-The labelled transition system for substitutions and cursor movement is defined as $(\ABT{}, Apc, \Rightarrow)$.
-This system describes the semantics of modifying the well-formed abt $a$ encapsulated by the cursor, by either substituting it with an operator or moving the cursor up or down the tree.
-Transitions are therefore of the form $a \ltrans{\pi} a'$, where $a, a' \in \ABT{}$ and $\pi \in Apc$.
-
-For substitution we define the transition rule for every label $\{o\} \in Apc$, which substitutes the abt currently encapuslated by the cursor with $o$.
-The general case of this rule is shown in \cref{fig:substitution_rules}.
-The side-condition ensures that we can only substitute operators of sort $s$ with abt's of sort $s$.
+
+The labelled transition system for substitutions and cursor movement
+is defined as $(\ABT{}, Apc, \Rightarrow)$. This system describes the
+semantics of modifying the well-formed abt $a$ encapsulated by the
+cursor, by either substituting it with an operator or moving the
+cursor up or down the tree. Transitions are therefore of the form $a
+\ltrans{\pi} a'$, where $a, a' \in \ABT{}$ and $\pi \in Apc$.
+
+For substitution we define the transition rule for every label $\{o\}
+\in Apc$, which substitutes the abt currently encapuslated by the
+cursor with $o$. The general case of this rule is shown in
+\cref{fig:substitution_rules}. The side-condition ensures that we can
+only substitute operators of sort $s$ with abt's of sort $s$.
+
 \federeboks{c}{
@@
 
-For cursor movement we define two transition rules "child  $i$" and "parent" for every operator $o$ of arity $(\Vec{s_1}.s_1, \dots,$ $ \Vec{s_n}.s_n)s$ and for every $1 \leq i \leq n$.
-These rules, shown in \cref{fig:cursor_movement_rules}, facilitate cursor movement from the parent operator to child $i$ and from child $i$ back to the parent operator, respectively.
+For cursor movement we define two transition rules "child  $i$" and
+"parent" for every operator $o$ of arity $(\Vec{s_1}.s_1, \dots,$ $
+\Vec{s_n}.s_n)s$ and for every $1 \leq i \leq n$. These rules, shown
+in \cref{fig:cursor_movement_rules}, facilitate cursor movement from
+the parent operator to child $i$ and from child $i$ back to the parent
+operator, respectively.
+
 \federeboks{c}{
@@
 \begin{example}\label{ex:substitution_rules}
-    Continuing from \cref{ex:well_formed_trees}, we now define the semantics of the editor calculus for our simple language.
+    We now define the semantics of the editor calculus for our simple language.
 We do not show the transition rules for editor expressions, since these are equivalent to the ones presented in \cref{fig:editor_expression_rules}.
@@
 }{\caption{Satisfaction relation for modal operators}\label{fig:satisfaction_relation_modal}}
-The modal operator $@o$ is satisfied when the root of the abt encapsulated by the cursor is the operator $o$. Therefore we define the rule (At-op) for every operator $o$.
-
-The modal operator $\Diamond o$ is satisfied when the operator $o$ is anywhere within the abt encapsulated by the cursor.
-For the trivial case, when the root of the tree encapsulated by the cursor is $o$, we only define the one rule (Possibly-trivial).
-For the non-trivial case, where $o$ is in one of the subtrees $\hat{a}_i$, we define a rule (Possibly-i) for every argument of every operator $o$.
-
-The modal operator $\square o$ is satisfied when $o$ is somewhere in every subtree of the abt encapsulated by the cursor. To describe this, we define the rule (Necessity) for every operator $o$.
-
-\begin{example}\label{ex:modal_logic}
-    We finalize our specialized editor calculus from \cref{ex:substitution_rules}, by defining the satisfaction relation.
+The modality $@o$ holds for an abt $a$ if the cursor is at the root of
+$a$ and the root is the operator $o$.
+
+The modality $\Diamond o$ is satisfied for $a$ when the operator $o$ is
+anywhere within $a$. The modality
+$\square o$ holds for $a$ when $o$ is found in every subtree of the
+abt encapsulated by the cursor.
+
+% \begin{example}\label{ex:modal_logic}
+%     We finalize our specialized editor calculus from \cref{ex:substitution_rules}, by defining the satisfaction relation.
 The propositional connectives are defined as previously shown in \cref{fig:satisfaction_relation_connectives}.
@@
 Rules for the remaining operators $o$ are defined accordingly.
-    \federeboks{c}{
-        \infeshort{At-plus}
-            {}
-            {[\text{plus}(\hat{a}_1;\hat{a}_2)]|= @\text{plus} }  
-        \\\\
-        \infeshort{Pos1-plus}
-            {[\hat{a}_1] |= \Diamond o}
-            {[\text{plus}(\hat{a}_1; \hat{a}_2)] |= \Diamond o}
-        \\\\
-        \infeshort{Nec-plus}
-            {[\hat{a}_1] |= \Diamond o \quad [\hat{a}_2] |= \Diamond o}
-            {[\text{plus}(\hat{a}_1;\hat{a}_2)] |=\square o}
-        \\\\
-        \infeshort{Pos2-plus}
-            {[\hat{a}_2] |= \Diamond o}
-            {[\text{plus}(\hat{a}_1; \hat{a}_2)] |= \Diamond o}
-    }{\caption{Satisfaction relation for modal operators on $plus$}\label{fig:example_satisfaction_relation}}
-\end{example}
+%     \federeboks{c}{
+%         \infeshort{At-plus}
+%             {}
+%             {[\text{plus}(\hat{a}_1;\hat{a}_2)]|= @\text{plus} }  
+%         \\\\
+%         \infeshort{Pos1-plus}
+%             {[\hat{a}_1] |= \Diamond o}
+%             {[\text{plus}(\hat{a}_1; \hat{a}_2)] |= \Diamond o}
+%         \\\\
+%         \infeshort{Nec-plus}
+%             {[\hat{a}_1] |= \Diamond o \quad [\hat{a}_2] |= \Diamond o}
+%             {[\text{plus}(\hat{a}_1;\hat{a}_2)] |=\square o}
+%         \\\\
+%         \infeshort{Pos2-plus}
+%             {[\hat{a}_2] |= \Diamond o}
+%             {[\text{plus}(\hat{a}_1; \hat{a}_2)] |= \Diamond o}
+%     }{\caption{Satisfaction relation for modal operators on $plus$}\label{fig:example_satisfaction_relation}}
+% \end{example}
+
+\subsection{Guarantees from the original editor calculus}
+
+The original editor calculus of \cite{type_safe_structure_editor} is equipped
+with a type system that types configurations relative to a path in the
+ast of the configuration. Judgments are therefore of the form $p, \Gamma_e \vdash
+\lr{E,a} : \textsf{ok}$, where $p$ is a path in the ast $a$ and
+$\Gamma_e$ is a type environment.
+
+The type system of the original calculus guarantees that a well-typed
+configuration remains well-typed after a transition and that a
+well-typed configuration will not get stuck. Here,
+$(p,\Gamma_e)\downarrow_\alpha$ denotes that the path and the type
+environment are updated because of the action $\alpha$. These
+guarantees can be expressed as the following theorem.
+
+\begin{theorem}[\cite{type_safe_structure_editor}]\label{thm:typesafe}
+  If $p, \Gamma_e \vdash \lr{E,a} : \textsf{ok}$ then
+  \begin{enumerate}
+  \item if $\lr{E,a} \ltrans{\alpha} \lr{E',a'}$, then
+    $(p,\Gamma_e)\downarrow_\alpha \vdash \lr{E',a'} : \textsf{ok}$
+
+  \item $p, \Gamma_e \vdash \lr{E,a} \neg\to \textsf{wrong}$
+  \end{enumerate}
+\end{theorem}
 
@@
 
+\subsection{A collection of lambda calculi}
+
 To encode the abstract binding trees, the simply typed lambda calculus
-suffices.
-
-In order to encode the generalized editor calculus using the simply
-typed $\lambda$-calculus (\stlc) as the base calculus, we extend the
-latter with new constructs.
-
-To encode the contexts, we introduce pairs. To encode atomic prefix
+suffices, so for an $a \in \ABT{}$ we have $\encode{a} \in \stlc$.
+
+In order to encode abts with cursors and expressions of the
+generalized editor calculus, we extend \stlc with more constructs.
+
+To represent contexts, we introduce pairs. To encode atomic prefix
 commands we further extend the calculus with pattern matching
@@
        \hline\hline
-       \ Abstract Binding Trees & \ Simply typed lambda calculus, \stlc \\
+       \ Abstract Binding Trees & \ Simply typed lambda calculus with pairs, \stlc \\
        \hline
-       \ Atomic Prefix Commands & \ \stlc with pattern matching, $\lambda_{->,p}$\\
+       \ Atomic Prefix Commands & \ \stlc with pattern matching $\lambda_{->,p}$\\
        \hline
@@
 
- All of the lambda calculi that we consider in the following have a
+ All of the lambda calculi that we consider have a
  standard reduction semantics with reductions of the form $e \to
- e'$. Moreover, every calculus has a type systems with type judgements
- of the form $\Gamma \vdash e : \tau$, where $\tau$ ranges over a set
- of types \Types\ and $\Gamma$ is a type environment
- $\Gamma : \Var \pra \Types$. The type systems all satisfy the subject
+ e'$.
+
+ Moreover, every calculus has a type systems with type judgements of
+ the form $\Gamma \vdash e : \tau$, where $\tau$ ranges over a set of
+ types \Types\ given by
+%
+\[  \tau ::= \tau_s \mid \tau \to \tau \mid \tau \times \tau  \]
+%
+where $\tau_s$ is a type for each sort $s \in \mathcal{S}$ and $\tau
+\to \tau$ and $\tau \times \tau$ are function and product types,
+respectively, and $\Gamma$ is a type environment
+ $\Gamma : \Var \pra \Types$.
+
+ The type systems all satisfy the subject
  reduction property.
@@
 \subsection{Abstract binding trees}
-To encode abstract binding trees we only need to use the simply typed lambda-calculus.
-We extend the lambda-calculus with term constants $o$ for every $o \in \OP$ excluding cursors, and the base types $s$ for every $s \in \SORT$.
+
+To encode abstract binding trees we only need to use the simply typed
+lambda-calculus where we have term constants $o$ for every $o \in \OP$
+excluding cursors, and the base types $s$ for every $s \in \SORT$.
+
 Typing rules for operators can be seen on \cref{fig:abt_typing_rules},
@@
 \begin{example}\label{ex:enc_abt_example}
-    Continuing from \cref{ex:modal_logic} we will now encode the operators in our small language.
-Following our singular rule for encoding abt's we get the resulting encoding seen on \cref{fig:enc_abt_example}, and the type for each operator on \cref{fig:abt_example_type}.
+    We can now encode the operators in our small language. Following
+    our singular rule for encoding abt's we get the resulting encoding
+    seen on \cref{fig:enc_abt_example}, and the type for each operator
+    on \cref{fig:abt_example_type}.
 
@@
     \encode{\pi.E} = \lambda C:\text{Ctx}.\encode{E} ( (\encode{\pi} \ C.1) \ , C.2) 
-    \\\\
+    \\[2mm]
     \encode{E_1 \ggg E_2}  = \lambda C:\text{Ctx}. \encode{E_2} (\encode{E_1} C)
-    \\\\
+    \\[2mm]
     \encode{\text{rec } x.E} = \text{fix } ( \lambda x: (\text{Ctx} \to \text{Ctx}).\encode{E}) 
-    \\\\
+    \\[2mm]
     \encode{\phi => E_1 | E_2} = \lambda C:\text{Ctx}.\text{match} \ (\encode{\phi} \ C.1) \\
         | \ \top ->  \encode{E_1} \ C 
-    \\
+    \\[2mm]
     | \  \bot ->  \encode{E_2} \ C 
-    \\\\
-    \encode{\lr{E, C[a']}} = \encode{E} \  (\encode{a'}, \encode{C})
-        \\\\
+    \\[2mm]
     \encode{\text{nil}} = \lambda C:\text{Ctx}.C \quad \quad    \encode{x} = x
@@
 }{
-    \caption{Encoding of editor expressions and context configuration.}
+    \caption{Encoding editor expressions}
     \label{fig:edt_enc}
@@
 
-\subsection{Soundness of the encoding}
+\subsection{Encoding configurations}
+
+It is now straightforward to encode configurations as follows.
+%
+\[ \encode{\lr{E, C[a]}} = \encode{E} \  (\encode{a}, \encode{C}) \]
+%
+Note that here $\encode{E} : T \times T_C \to T$ where $T$ is the type
+of the encoding of an abt and $T_C$ is the type of the encoding of a context.
+
+\section{Soundness of the encoding}
 
@@
 
+Because of Theorem \ref{thm:subred}, we now get that our encoding is
+type-safe.
+
+\begin{corollary}
+  If $\Gamma \vdash \encode{\lr{E,C[a]}} : t$ then, if $\lr{E,a}
+  \ltrans{} \lr{E',a'}$, we also have $\Gamma \vdash \encode{\lr{E',C[a']}} : t$.
+\end{corollary}
+
 \section{Conclusion and Further Work}
@@
 creation of a syntax-directed editor calculus for a specific abstract
-syntax. Subsequently, we encoded this editor calculus into an extended
-version of the lambda calculus, which incorporates pairs, recursion
-and pattern matching.  Further work on our generalized editor calculus
-could be to prove whether the encoding is complete. Otherwise, one
-could implement a redo-undo construct such as introduced in
-\cite{undo_Edit} in the generalized editor calculus. Another
-interesting construct to implement would be a copy-paste operation.
-
-%Furthermore, we have proven the soundness of the encoding of the generalized editor calculus into the extended lambda calculus.
-As a consequence, given that the type system of the extended lambda calculus is sound, it logically follows that our editor calculus also possesses a sound type system.
-
-We have identified several possibilities for expanding our work on the
-generalized editor calculus. First of all, the encoding should be
-proved complete. Secondly,
+syntax, and describe how to encode it into a monomorphic typed lambda
+calculus with pairs, recursion and pattern matching. The encoding is
+sound, and as a consequence, given that the type system of the
+extended lambda calculus is sound, it follows that our editor calculus
+is typesafe.
+
+Further work on our generalized editor calculus is to prove whether
+the encoding is complete.  An important next step is to incorporate a
+copy-paste operation. Here, the question arises as to how to best deal
+with pasting terms that include free variables. In some settings, when
+editing code, this will introduce name clashes, but in other settings
+the free variables should intentionally be captured. A likely solution
+is therefore to have two different paste operations corresponding to
+each of these situations.
 
@@
 \endinput
-%%
-%% End of file `sample-sigplan.tex'.
 

### 2023-08-25 Response to reviewers.tex
Create Response to reviewers.tex

\documentclass[anonymous,review]{acmart}
\input{preamble}

%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2023}
\acmYear{2023}
\acmDOI{XXXXXXX.XXXXXXX}

\acmConference[GPCE 2023]{ }{July 07--07,
  2023}{Woodstock, NY}

\acmPrice{15.00}
\acmISBN{978-1-4503-XXXX-X/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% For managing citations, it is recommended to use bibliography
%% files in BibTeX format.
%%
%% You can then either use BibTeX with the ACM-Reference-Format style,
%% or BibLaTeX with the acmnumeric or acmauthoryear sytles, that include
%% support for advanced citation of software artefact from the
%% biblatex-software package, also separately available on CTAN.
%%
%% Look at the sample-*-biblatex.tex files for templates showcasing
%% the biblatex styles.
%%

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{\emph{Response to reviewers:} A type-safe generalized editor calculus (Short Paper)}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.

%%\orcid{1234-5678-9012}

%\author{Andreas Tor Mortensen}
%\email{atmo20@student.aau.dk}

%\author{Benjamin Bennetzen}
%\email{bbenne20@student.aau.dk}

%\author{Nikolaj Rossander Kristensen}
%\email{nrkr20@student.aau.dk}

%\author{Peter Buus Steffensen}
%%\authornotemark[1]
%\email{psteff19@student.aau.dk
%}

%\affiliation{%
%  \institution{Department of Computer Science at Aalborg University}
%  \streetaddress{Selma Lagerlöfsvej 300}
%  \city{Aalborg}
%  \country{Denmark}
%  \postcode{9220}
%}

%\author{Hans Huttel}
%\email{hans@cs.aau.dk}
%\affiliation{%
%  \institution{Department of Computer Science at Aalborg University}
%  \streetaddress{Selma Lagerlöfsvej 300}
 % \city{Aalborg}
 % \country{Denmark}
 % \postcode{9220}
%}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{AAU Bachelor}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011050.10011023</concept_id>
       <concept_desc>Software and its engineering~Specialized application languages</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10010124.10010131</concept_id>
       <concept_desc>Theory of computation~Program semantics</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10010124.10010131.10010134</concept_id>
       <concept_desc>Theory of computation~Operational semantics</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Specialized application languages}
\ccsdesc[500]{Theory of computation~Program semantics}
\ccsdesc[500]{Theory of computation~Operational semantics}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{Syntax-directed editor, Generalisation, lambda calculus}
%% A "teaser" image appears between the author and affiliation
%% information and the body of the document, and typically spans the
%% page.
%\begin{teaserfigure}
 % \includegraphics[width=\textwidth]{Images/windmill.jpg}
 % \caption{Seattle Mariners at Spring Training, 2010.}
%  \Description{Enjoying the baseball game from the third-base
%  seats. Ichiro Suzuki preparing to bat.}
%  \label{fig:teaser}
%\end{teaserfigure}

%\received{20 February 2007}
%\received[revised]{12 March 2009}
%\received[accepted]{5 June 2009}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

This response consists of

\begin{itemize}
\item Full responses to the comments and questions from the reviewers.
\item The full soundness proof, as requested by Reviewer 2.
\end{itemize}

\newcommand{\reviewer}[1]{\textsl{#1}}
\newcommand{\response}[1]{\paragraph{\textbf{Response:}} #1}

\section{Response}

\subsection{Review \#58A}

\begin{itemize}
\item \reviewer{L76: I see that Andersen et al. had to extend an
    existing editor calculus with let expressions, but I don't see how
    it is related to the previous sentence. Did they write the type
    safety proof from scratch?}
  \response{Some of the type safety proof had
  to be redone -- because of the presence of let-expressions
  in the underlying language and because of parametric
  polymorphism. The addition of a copy/paste primitive in the editor
  calculus was the cause of other revisions. }
\item \reviewer{L80: I could not figure out what "original work"
    refers to.}
  \response{This refers to the Cornell Program Synthesizer mentioned
    in lines 32-33.}
\item \reviewer{L91: I do not see any discussion of higher-order
    abstract syntax in the subsequent sections.}
\response{Indeed. We will rephrase this -- we simply take inspiration from
  this approach to representing abstract syntax.}
\item \reviewer{Example 2.1: Please explain what $s$ and $e$
    represent, and what the notation $e.s$ means. Also, I think the
    term "sort" is not used in the standard way.}
  \response{$s$ is a metavariable for "scoped expressions'' and $e$ is
  a metavariable for "simple expressions''. We will add this
  explanation. The notation $(e)s$ is the arity of the term constructor
  for scoped expressions that are simple expressions: it takes an $e$
  and creates an $s$.

Our presentation of abstract syntax and the terminology and notation
(includion the notions of sorts and
arities) closely follows that of Harper \cite{harper_foundations}, as
mentioned at the start of Section 2. We can elaborate on this approach
at that point in the text.}
\item \reviewer{Section 2.2: Please explain what each syntactic form
    represents.}
  \response{Editor expressions $E$ represent operations on abstract
    binding trees $a$. This has already been explained in
    Subsection 2.1.1. But indeed a short explanation of editor
    expressions in Section 2.1.1 is called for. It is:

    \begin{itemize}
    \item $\pi$ denote navigation actions that lets the cursor move
      down the abt ($\text{child} \; n$ and $\text{parent} $) and
      substitution actions that replace the subtree pointed to by the
      cursor by the syntactic operator $o$.
    \item Expressions $\pi.E$ denote performing $\pi$ and continuing
      as $E$. $\phi \Rightarrow E_1|E_2$ is a conditional expression whose
      continuation depends on the condition $\phi$. $E_1 \ggg E_2$ is
      sequential composition. $\text{rec}\;x.E$ is a recursive
      expression; occurrences of $x$ in $E$ are recursive
      calls. $\text{nil}$ is the empty editor expression.
    \end{itemize}

    We will add this explanation to Section 2.1.1.}

\item \reviewer{L174: It is not very clear what the authors mean by
    "encapsulated by the cursor".}
  \response{The cursor points to the root a subtree of the abt being
    edited, so the abt $a$ can be written as a cursor context $C[a_1]$
    whenever the cursor is at the root of the subtree $a_1$. See also
    the end of Section 2.1.2.}
\end{itemize}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}
\endinput

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
